---
title: "maps_design"
author: "Magali Weissgerber"
date: "2026-02-04"
output: html_document
---

## Loading libraries

```{r}

if (!requireNamespace("renv", quietly = TRUE)) install.packages("renv")
renv::activate()
renv::restore()

install.packages('sf') # Provides functions to read, write, and manipulate spatial vector data (points, lines, polygons).
library(sf)

install.packages('data.table') # Provides functions for efficient manipulation and processing of large tabular datasets.
library(data.table)

install.packages('terra') # Used for handling, analyzing, and visualizing spatial raster data.
library(terra)

install.packages('ravetools') # for the function convolve_image
library(ravetools)

install.packages('dplyr') # Provides functions for common data manipulation tasks such as filtering, selecting, and summarizing data.
library(dplyr)

# For plotting

install.packages('ggplot2') # Used to create data visualizations using the grammar of graphics framework.
library(ggplot2)

install.packages("cowplot") # Provides functions for aligning and combining multiple ggplot2 plots.
library(cowplot)

```

## Prepare helpful necessary layers
- A reference raster
- A mask with Europe's borders
- Preprocessed (masked) human population density raster
- Europe borders and coastlines

```{r}

# We use a raster of human population density, it can be downloaded here https://ec.europa.eu/eurostat/web/gisco/geodata/population-distribution/population-grids
# We use it as a reference raster for resolution, crs, and extent

# Import population raster not distinguishing inland (should be 0) and oceans (should be NA)
# Step 1: Load the population raster and coastline shapefile
raster_1km_human_population_path <- "S:/Magali_Weissgerber/human_population_scratch/JRC_1K_POP_2018.tif" # Read the 1km resolution raster
raster_1km_human_population <- rast(raster_1km_human_population_path)
plot(raster_1km_human_population)

# It is not distinguishing different type of absence of popupation: inland (should be 0) and oceans (should be NA)
# So we are using a file with the coastlines to make this disctinction

# Step 2: Load coastline shapefile
# Import Europe's borders (download here #### this one we can put online ####)
inland_shapefile <- st_read("S:/Magali_Weissgerber/Europe frontiers_scratch/NUT3_from_rewilding_folder/Europe_limit_from_nut3_dissolve.shp")
plot(inland_shapefile)

# Step 3: check the objects
# Check the type of the raster object
class(raster_1km_human_population)
# Check if the raster object is valid
print(raster_1km_human_population)
# Ensure the CRS is consistent
crs(raster_1km_human_population) <- "EPSG:3035"
# Check and reproject the shapefile to match the CRS of the raster
if (crs(inland_shapefile) != crs(raster_1km_human_population)) {
  inland_shapefile <- st_transform(inland_shapefile, crs(raster_1km_human_population))
}

# Step 4: Create a mask from the coastline shapefile
# Rasterize the coastline shapefile to create a mask
inland_raster <- rasterize(vect(inland_shapefile), raster_1km_human_population, field = 1, background = NA)
plot(inland_raster)

# Step 5: Preprocess the population raster
# Replace inland NA values with 0
population_preprocessed <- raster_1km_human_population
values(population_preprocessed)[is.na(values(population_preprocessed)) & !is.na(values(inland_raster))] <- 0
plot(population_preprocessed)

#Rename
raster_1km_human_population <- population_preprocessed
#Define the reference raster
ref_raster <- raster_1km_human_population

#Prepare also country borders and coastlines for mapping
country_borders <- st_read("I:/biocon/Projects/RewildingMAP/DataSources/ReferenceData/ref-countries-2016-01m/CNTR_BN_01M_2016_3035_INLAND/CNTR_BN_01M_2016_3035_INLAND.shp")
coastlines <- st_read("I:/biocon/Projects/RewildingMAP/DataSources/Europe_coastline_shapefile/Europe_coastline.shp")

```

## I. Upload and prepare the data

# 1 Climate change mitigation

Current = disturbance regime highly natural

```{r}

# Intensity of use with threshold

# Path to the HANPP raster (HANPP in percent of potential NPP for the year 2006)
hanpp_path <- "I:/biocon/Projects/RewildingMAP/Consolidated/BASE_LAYERS/hanpp2006gapfilled2000_1k/hanpp06gapf00intNodata.tif"
# Load the raster
hanpp_raster <- rast(hanpp_path)
# Plot to check
plot(hanpp_raster, main = "HANPP 2006 Gapfilled (1km)")
# Invert and re-scale (100 - value) / 100 (we want high values to indicate natural areas and the % to be between 0 and 1)
hanpp_inverted <- (100 - hanpp_raster) / 100
# Checks
crs(hanpp_inverted) == crs(ref_raster)             # Coordinate system
res(hanpp_inverted) == res(ref_raster)             # Resolution
ext(hanpp_inverted) == ext(ref_raster)             # Extent
hanpp_proj <- project(hanpp_inverted, ref_raster)
plot(hanpp_proj)
hanpp_aligned <- terra::resample(hanpp_proj, ref_raster, method = "near")  # use "bilinear" for continuous data
# Plot to check result
plot(hanpp_aligned, main = "Inverted HANPP/NPP (scaled 0–1)")
# Potential expansion with a combined probability of at least 0.8 (threshold)
hanpp_aligned_threshold_8 <- classify(hanpp_aligned, matrix(c(-Inf, 0.8, 0, 0.8, Inf, 1), ncol = 3, byrow = TRUE))
# Plot to check
plot(hanpp_aligned_threshold_8, main = "Binary naturalness (≥0.8 = 1)")
# Masked so only areas of interest are kept
hanpp_masked <- mask(hanpp_aligned_threshold_8, inland_raster)
plot(hanpp_masked)

```

Potential = abandonment

```{r}

# Abandonment

abandon <- terra::rast('I:\\biocon\\Magali_Weissgerber\\wildE_framework_240912\\Maps\\20250106_lq\\data\\rewilding.tif')
# (1 = 1 scenario of abandonment is happening, 4 = four scenarios of abandonment are happening)
plot(abandon)
# Classify abandonment to present/absent
rcl <- matrix(c(1,1,
                2,1,
                3,1,
                4,1,
                0,NA,
                9,NA,
                255, NA), ncol=2, byrow = T)
aban_simple <- terra::classify(abandon, rcl)
plot(aban_simple)
# Check crs, red and ext
crs(aban_simple) == crs(ref_raster)             # Coordinate system
res(aban_simple) == res(ref_raster)             # Resolution
ext(aban_simple) == ext(ref_raster)             # Extent
aban_simple_proj <- project(aban_simple, ref_raster)
plot(aban_simple_proj)
aban_simple_aligned <- terra::resample(aban_simple_proj, ref_raster, method = "near")  # use "bilinear" for continuous data
plot(aban_simple_aligned)
# Masked so only areas of interest are kept
abandonment_masked <- ifel(
  is.na(inland_raster), 
  NA,                      # outside → NA
  ifel(is.na(aban_simple_aligned), 0, aban_simple_aligned)  # inside → NA → 0
)
plot(abandonment_masked)

```

Benefits = potential carbons stock

```{r}

# Potential carbon (Walker 2022), data are available here https://doi.org/10.7910/DVN/DSDDQK 

pot_carbon <- terra::rast('S:\\Magali_Weissgerber\\HANPP_scratch\\Unrealised_carbon_stock_Walker2022\\Base_Pot_AGB_BGB_SOC_MgCha_500m.tif')
crs(pot_carbon) == crs(ref_raster)             # Coordinate system
res(pot_carbon) == res(ref_raster)             # Resolution
ext(pot_carbon) == ext(ref_raster)             # Extent
pot_carbon_proj <- project(pot_carbon, ref_raster)
# Masked so only areas of interest are kept
pot_carbon_masked <- mask(pot_carbon_proj, inland_raster)
plot(pot_carbon_masked)

```

# 2 Climate change adaptation

Current = high effectivement mesh size

```{r}

# Connectivity

# Load the file
# Define import directory
# Define the directory path
directory_anth_meff <- "I:\\biocon\\Emmanuel_Oceguera\\projects\\2025_10_WildE_fragmentation_magali\\fragmentation\\ESACCI_LULC\\output\\natural_areas\\10km_radius"
# List files for 2022 with .tif or .tiff extension
meff_2022_file <- list.files(
  path = directory_anth_meff,
  pattern = "2022.*\\.tif[f]?$",  # matches .tif or .tiff
  full.names = TRUE
)
# Import the raster
meff_2022 <- rast(meff_2022_file)
# Checks
crs(meff_2022) == crs(ref_raster)             # Coordinate system
res(meff_2022) == res(ref_raster)             # Resolution
ext(meff_2022) == ext(ref_raster)             # Extent
meff_2022_proj <- project(meff_2022, ref_raster)
plot(meff_2022_proj)
# Keep higher 80%
# Reclassify: values >= 80 → 1, else → 0
meff_2022_binary_T8 <- ifel(meff_2022_proj >= 80, 1, 0)
plot(meff_2022_binary_T8)
# Masked so only areas of interest are kept
high_meff_masked <- mask(meff_2022_binary_T8, inland_raster)
plot(high_meff_masked)

```

Potential = abandonment (cf Climate change mitigation)

Benefits = climate velocity

```{r}

# Climate velocity

cv <- terra::rast('S:\\Magali_Weissgerber\\Climate velocity\\CV_data_cropped_to_europe.tif')
plot(cv)
crs(cv) == crs(ref_raster)             # Coordinate system
res(cv) == res(ref_raster)             # Resolution
ext(cv) == ext(ref_raster)             # Extent
cv_proj <- project(cv, ref_raster)
plot(cv_proj)
cv_aligned <- terra::resample(cv_proj, ref_raster, method = "near")  # use "bilinear" for continuous 
plot(cv_aligned)
# Masked so only areas of interest are kept
cv_masked <- mask(cv_aligned, inland_raster)
plot(cv_masked)

```

# 3 Social benefits and risk

# 3.1 Determining the expansion probability of large mammals

We are using the wildlife maps from the Rewild Europe Report (Fernández, N., Torres, A., Wolf, F., Quintero, L., & Pereira, H. M. (2020). Boosting Ecological Restoration for a Wilder Europe. Halle-Wittenberg: German Centre for Integrative Biodiversity Research (iDiv) and Martin-Luther-Universität.)

They include Alpine ibex, Brown bear, Chamois, Elk, European bison, European lynx, European wolf, Iberian ibex, Iberian lynx, Pyrenean chamois, Red deer, Reindeer, Roe deer, Wild boar, Wild goat.
We will focus on Brown bear, Elk, European bison, European wolf, European lynx, Iberian lynx, Iberian ibex, Alpine ibex.

We want to know the probability that cells not containing the focus species will gain that species. Here are the different steps.
For each species
1) Occurrences are coded as 0 = species absent, 1 = species present and 9 = species maybe present. We are changing all the 9 into 0.
2) Get the distance of every 0 to the closest 1
3) Calculate probability of colonization (form K/e^-xdistance)
Then
4) Sum the probability (Optional)

A parallel processing was added to rasterize the data directly and change the resolution from 10km to 1km, and calculate distance then.

Import and check data; change to binary

```{r}

SPECIES_OCC_initial_fishnet <-  st_read("S:/Magali_Weissgerber/Species_dist_occ_scratch/SP_OCC_HIST_2019report/SP_OCCU_190207.shp")

head(SPECIES_OCC_initial_fishnet)
names(SPECIES_OCC_initial_fishnet)

ggplot(SPECIES_OCC_initial_fishnet) + 
  geom_sf(aes(fill = M_CANLUP), color = NA) +
  scale_fill_viridis_c(option = "magma",begin = 0.2)

SPECIES_OCC_binary_fishnet <- SPECIES_OCC_initial_fishnet #new dataset to be modified
  
# Change 9 into 0
  for(j in 4:20)
    set(SPECIES_OCC_binary_fishnet, i = which(SPECIES_OCC_binary_fishnet[[j]] == 9), j = j, value = 0)

#Plot one species to check
ggplot(SPECIES_OCC_binary_fishnet) + 
  geom_sf(aes(fill = M_CANLUP), color = NA) +
  scale_fill_viridis_c(option = "magma",begin = 0.2)

```

Rasterize, create distance raster and calculate probability of expansion

Calculate probability of colonization (form K/e^-xdistance) (when distance = 0 it means the species is already there so probability is also 0). 
Initially we used 1/exp((log(2)/50) distance) so to have 50% of probability of comeback at 50km. But based on Canis lupus dispersal values we apply rather 1/exp((log(2)/60) distance) (as the cells of the mammals distribution are 10x10km the probability max is 0.89)

First, we rasterize the shapefile. It becomes a stack raster with the 8 species of interest

```{r}

# Define the extent and resolution for the raster
extent_raster <- terra::ext(SPECIES_OCC_binary_fishnet)
res <- 10000  # Define your desired resolution
# Create an empty raster with the defined extent and resolution
r <- terra::rast(extent_raster, resolution = res, crs = crs(SPECIES_OCC_binary_fishnet))
# List of columns to rasterize
columns_to_rasterize <- c("M_CANLUP", "M_ALCALC", "M_BISBON", "M_LYNLYN", "M_URSARC", "M_LYNPAR", "M_CAPIBE", "M_CAPPYR")
# Initialize an empty list to store rasters
rasters <- list()
# Loop through each column and rasterize
for (col in columns_to_rasterize) {
    r <- rasterize(SPECIES_OCC_binary_fishnet, r, field = col)
    rasters[[col]] <- r
}
# Stack all rasters
raster_10km_wildlife_occupancy_stack <- rast(rasters)
plot(raster_10km_wildlife_occupancy_stack)

# Change the resolution of the wildlife raster 

# Check the original raster resolutions
print(res(raster_10km_wildlife_occupancy_stack))
print(res(raster_1km_human_population))

# Resample the 10km raster to 1km resolution using nearest neighbor method
raster_1km_wildlife_occupancy_stack <- resample(
  raster_10km_wildlife_occupancy_stack,
  raster_1km_human_population,
  method = "near"
)

# Check the resampled raster resolution
print(res(raster_1km_wildlife_occupancy_stack))

# Plot raster
plot(raster_1km_wildlife_occupancy_stack)

# Sum the occupancy (for illustrative maps)
raster_1km_wildlife_occupancy_stack$M_sum <- app(raster_1km_wildlife_occupancy_stack, sum)
plot(raster_1km_wildlife_occupancy_stack)

```

Now we produce the distance raster for each species (depending on processing power it can take multiple hours per species to process)

```{r}
# 1. Set output directory here

output_directory <- "S:/Magali_Weissgerber/Species_dist_occ_scratch/species_distance_260213"

# 2. Create list of species rasters

species_list <- list(
  CANLUP = raster_1km_wildlife_occupancy_stack$M_CANLUP,
  ALCALC = raster_1km_wildlife_occupancy_stack$M_ALCALC,
  BISBON = raster_1km_wildlife_occupancy_stack$M_BISBON,
  LYNLYN = raster_1km_wildlife_occupancy_stack$M_LYNLYN,
  URSARC = raster_1km_wildlife_occupancy_stack$M_URSARC,
  LYNPAR = raster_1km_wildlife_occupancy_stack$M_LYNPAR,
  CAPIBE = raster_1km_wildlife_occupancy_stack$M_CAPIBE,
  CAPPYR = raster_1km_wildlife_occupancy_stack$M_CAPPYR
)

# 3. Processing function

process_and_export_species <- function(r, species_name, output_dir) {

  tryCatch({

    message("=======================================")
    message("Processing species: ", species_name)

    # Step 1: Clean raster
    r_clean <- classify(r, matrix(c(NaN, NA), ncol = 2, byrow = TRUE))
    r_clean <- mask(r_clean, r_clean)

    # Step 2: Extract presence cells
    pts <- as.points(r_clean, values = TRUE, na.rm = TRUE)
    pts <- pts[pts[[1]] == 1, ]

    if (nrow(pts) == 0) {
      stop("No presence cells found.")
    }

    # Step 3: Compute distance raster
    dist_r <- distance(r_clean, pts, unit = "km")

    # Step 4: Export
    output_file <- file.path(
      output_dir,
      paste0("dist_", species_name, ".tif")
    )

    writeRaster(dist_r, output_file, overwrite = TRUE)

    message("Finished and exported: ", output_file)

    return(TRUE)

  }, error = function(e) {

    message("ERROR in species ", species_name, ": ", e$message)
    return(FALSE)

  })
}

# 4. Loop over all species

results_status <- list()

for (sp in names(species_list)) {

  results_status[[sp]] <- process_and_export_species(
    r = species_list[[sp]],
    species_name = sp,
    output_dir = output_directory
  )

}

# 5. Summary of processing

message("=======================================")
message("Processing summary:")
print(results_status)
```

# Next, import the distance raster

```{r}
# 1. List species names

species <- c("CANLUP", "ALCALC", "BISBON", "LYNLYN", "URSARC", "LYNPAR", "CAPIBE", "CAPPYR")
# Directory where files were exported
input_dir <- "S:/Magali_Weissgerber/Species_dist_occ_scratch/species_distance_260213"

# 2. Create a function: import + plot + mask + return

import_and_mask_species <- function(species, input_dir, inland_raster) {
  
  # Find the correct file automatically (matching species name)
  file_pattern <- paste0("dist_", species, ".tif")
  file_path <- list.files(input_dir, pattern = file_pattern, full.names = TRUE)
  
  if (length(file_path) == 0) {
    stop("No file found for species: ", species)
  }
  
  message("Importing raster for: ", species)
  
  # Import
  r <- rast(file_path)
  plot(r, main = paste("Imported Distance Raster:", species))
  
  # Mask
  r_masked <- mask(r, inland_raster)
  plot(r_masked, main = paste("Masked Distance Raster:", species))
  
  return(r_masked)
}

# 3. Apply this function to all species automatically

masked_rasters <- lapply(
  species,
  import_and_mask_species,
  input_dir = input_dir,
  inland_raster = inland_raster
)

# The list will be named for convenience:
names(masked_rasters) <- paste0("dist_", species)

# Check results

masked_rasters$dist_LYNPAR
masked_rasters$dist_CAPIBE
masked_rasters$dist_CAPPYR
```

Then, calculate probability for expansion (based on the distance, which is the minimum distance of each 0 cell (=absence) to)

```{r}
# Step 1: Define the transformation function
probability_function <- function(distance) {
  ifelse(distance == 0, 
         0, 
         1 / exp((log(2) / 60) * distance))
}

# Apply the function to each raster in the stack
species_probability_stack <- lapply(
  masked_rasters,
  \(r) terra::app(r, probability_function)
)

# Rename the rasters: replace "dist_" with "P_" in layer names (the distribution fields are called dist_CANLUP, etc)
old_names <- names(masked_rasters)               # Get original names
new_names <- gsub("^dist_", "P_", old_names)       # Replace "dist_" and add "P_" (for "probability")
names(species_probability_stack) <- new_names   # Assign new names to result_stack

# Convert the list of rasters into a multi-layer SpatRaster
species_probability_stack <- rast(species_probability_stack)

# Calculate the sum of the 8 species rasters
sum_raster <- sum(species_probability_stack)

# Rename the sum raster
names(sum_raster) <- "P_SUM"

# Add the sum raster to the stack
species_probability_stack_2 <- c(species_probability_stack, sum_raster)

# Plot results
plot(species_probability_stack_2)

#Export
#output_file_Proba <- ("S:/Magali_Weissgerber/Species raster for social benefit of rewilding/SP_expansion_probability_250106.tif")
#writeRaster(species_probability_stack_2, output_file_Proba, overwrite=TRUE)

```

## 3.2 - determining social benefit

We look at people's willingness-To-Travel (WTT) to see wildlife.
WTT is distributed among people, we use this distribution to determine the relation between distance to a cell and the share of people in the surrounding cells that will travel. Use the complementary lognormal CDF function (1 - CDF) and the density of people in a convolution to determine the amount of people going to one cell.

Prepare rasters

```{r}

# The human population raster has been pre processes already 
plot(raster_1km_human_population)

# Check the rasters of mammals distribution and expansion probability

names(raster_1km_wildlife_occupancy_stack)
res(raster_1km_wildlife_occupancy_stack)
names(species_probability_stack_2)
res(species_probability_stack_2)

# Current occupancy + probability

# Prepare a raster with both current distribution and probability
raster_1km_wildlife_current_probability_stack <- c(raster_1km_wildlife_occupancy_stack, species_probability_stack_2)

# Check the raster resolution
print(res(raster_1km_wildlife_current_probability_stack))

#Plot raster
plot(raster_1km_wildlife_current_probability_stack)

#Sum the probability (for illustrative maps)
#raster_1km_wildlife_current_probability_stack$MP_sum <- app(raster_1km_wildlife_current_probability_stack, sum)
#plot(raster_1km_wildlife_current_probability_stack)
#Export
#output_file_MP <- ("S:/Magali_Weissgerber/Species raster for social benefit of rewilding/SP_OCCU_current_probability_241212.tif")
#writeRaster(raster_1km_wildlife_current_probability_stack, output_file_MP, overwrite=TRUE)

# We look at people's willingness-To-Travel to see wildlife Possibility. In case some areas should be excluded (e.g. rural areas) we add a step of creating a new raster with a threshold of population density (threshold can be applied now, the current version has no threshold)
raster_1km_human_population_threshold <- raster_1km_human_population

```

Calculate the number of people travelling based on the WTT

For large carnivores

```{r}

# Step 1: create the kernel matrix

# Step 1.1: Define matrix dimensions and create the distance matrix
nrows <- nrow(raster_1km_human_population_threshold)
ncols <- ncol(raster_1km_human_population_threshold)
x_coords <- matrix(rep(seq(1, ncols), each = nrows), nrow = nrows, ncol = ncols)
y_coords <- matrix(rep(seq(1, nrows), times = ncols), nrow = nrows, ncol = ncols)
center_x <- ncols / 2
center_y <- nrows / 2
distance_matrix_LC <- sqrt((x_coords - center_x)^2 + (y_coords - center_y)^2)

# Step 1.2: Define parameters for the lognormal kernel
mu <- -2.23846
sigma <- 1.000746

# Step 1.3: Create the lognormal kernel matrix
kernel_matrix_LC <- dlnorm(distance_matrix_LC, meanlog = mu, sdlog = sigma)

# Step 1.4: Check the dimensions and a small portion of the matrix
dim(kernel_matrix_LC)
print("Top-left corner of the matrix:")
print(kernel_matrix_LC[1:10, 1:10])  # Displaying the top-left 10x10 cells

#Step 2: create the population matrix (the ocean's 0 were masked and change to NAs previously because convolution was done in the spatial domain, with this faster method using fourier transform the NAs have to be set back to 0 for the matrix, yet I kept it that way)

#Step 2.1: create an identifcal raster and print basic information about the raster
raster_1km_human_population_threshold_2 <- raster_1km_human_population_threshold
print(raster_1km_human_population_threshold_2)
# Summary of the raster to check the range of values
summary(raster_1km_human_population_threshold_2)

#Step 2.2: Check the number of non-zero and NA values
num_zeros <- sum(values(raster_1km_human_population_threshold_2) == 0, na.rm = TRUE)
num_nas <- sum(is.na(values(raster_1km_human_population_threshold_2)))
num_non_zeros <- sum(values(raster_1km_human_population_threshold_2) > 0, na.rm = TRUE)
cat("Number of zero values:", num_zeros, "\n")
cat("Number of NA values:", num_nas, "\n")
cat("Number of non-zero values:", num_non_zeros, "\n")

#Step 2.3: Replace NA values with zeros
raster_1km_human_population_threshold_2[is.na(raster_1km_human_population_threshold_2)] <- 0
plot(raster_1km_human_population_threshold_2)
# Check if NA values are replaced
na_count_after <- sum(is.na(values(raster_1km_human_population_threshold_2)))
print(paste("Number of NA values after replacement:", na_count_after))

#Step 2.4: Get the number of rows and columns from the original raster
nrows <- nrow(raster_1km_human_population_threshold_2)
ncols <- ncol(raster_1km_human_population_threshold_2)

#Step 2.5: Convert the raster to a matrix and reshape it
population_matrix_LC <- matrix(values(raster_1km_human_population_threshold_2), 
                            nrow = nrows, ncol = ncols, byrow = TRUE)
# Check the dimensions and a small portion of the matrix
dim(population_matrix_LC)
print("Top-left corner of the matrix:")
print(population_matrix_LC[1:10, 1:10])  # Displaying the top-left 10x10 cells

# Step 3: Use the 'Fast-Fourier' transform to compute the convolution
convolved_matrix_LC <- convolve_image(population_matrix_LC, kernel_matrix_LC)

# Step 4: convert the matrix to raster

# Step 4.1: Convert the convolved matrix back to raster format using terra's rast function
convolved_raster_LC <- rast(nrows = nrow(raster_1km_human_population_threshold_2),
                                  ncols = ncol(raster_1km_human_population_threshold_2),
                                  xmin = xmin(raster_1km_human_population_threshold_2),
                                  xmax = xmax(raster_1km_human_population_threshold_2),
                                  ymin = ymin(raster_1km_human_population_threshold_2),
                                  ymax = ymax(raster_1km_human_population_threshold_2))

# Step 4.2: Assign the values of the convolved matrix back to the raster
values(convolved_raster_LC) <- convolved_matrix_LC

# Step 4.3: fix crs, extent and resolution
res(raster_1km_human_population_threshold_2)
res(convolved_raster_LC)
crs(raster_1km_human_population_threshold_2)
crs(convolved_raster_LC)
ext(raster_1km_human_population_threshold_2)
ext(convolved_raster_LC)

# Step 4.3: fix crs
crs(convolved_raster_LC) <- crs(raster_1km_human_population_threshold_2)

# Step 5: Plot results
plot(raster_1km_human_population_threshold_2, main = "Original Population Raster")
plot(kernel_matrix_LC, main = "Kernel Raster")
plot(convolved_raster_LC, main = "Convolved Raster with Identity Kernel")

# Step 6: mask the ocean

# Step 6.1: import the inland shapefile
inland_shapefile <- st_read("S:/Magali_Weissgerber/Europe frontiers_scratch/NUT3_from_rewilding_folder/Europe_limit_from_nut3_dissolve.shp")

# Step 6.2: mask the output raster
convolved_raster_masked_LC <- mask(convolved_raster_LC, inland_raster)
plot(convolved_raster_masked_LC)

# Compute quantile breaks (exclude NA values)
q_breaks <- quantile(values(convolved_raster_masked_LC), 
                     probs = seq(0, 1, length.out = 11), 
                     na.rm = TRUE)

# Plot with 10 quantile levels
plot(convolved_raster_masked_LC, 
     breaks = q_breaks, 
     col = hcl.colors(10, "YlOrRd", rev = TRUE), 
     main = "Convolved Raster (Quantile Breaks) large carnivores")

# Step 7: export
#output_file_WTT_masked <- ("S:/Magali_Weissgerber/Species raster for social benefit of rewilding/WTT_LC_raster_masked_240925.tif")
#writeRaster(convolved_raster_masked_LC, output_file_WTT_masked, overwrite=TRUE)

```

For large herbivores

```{r}

# Step 1: create the kernel matrix

# Step 1.1: Define matrix dimensions and create the distance matrix
nrows <- nrow(raster_1km_human_population_threshold)
ncols <- ncol(raster_1km_human_population_threshold)
x_coords <- matrix(rep(seq(1, ncols), each = nrows), nrow = nrows, ncol = ncols)
y_coords <- matrix(rep(seq(1, nrows), times = ncols), nrow = nrows, ncol = ncols)
center_x <- ncols / 2
center_y <- nrows / 2
distance_matrix_LH <- sqrt((x_coords - center_x)^2 + (y_coords - center_y)^2)

# Step 1.2: Define parameters for the lognormal kernel
mu <- -1.50609
sigma <- 0.572612

# Step 1.3: Create the lognormal kernel matrix
kernel_matrix_LH <- dlnorm(distance_matrix_LH, meanlog = mu, sdlog = sigma)

# Step 1.4: Check the dimensions and a small portion of the matrix
dim(kernel_matrix_LH)
print("Top-left corner of the matrix:")
print(kernel_matrix_LH[1:10, 1:10])  # Displaying the top-left 10x10 cells

#Step 2: create the population matrix

#Step 2.1: create an identifcal raster and print basic information about the raster
raster_1km_human_population_threshold_3 <- raster_1km_human_population_threshold
print(raster_1km_human_population_threshold_3)
# Summary of the raster to check the range of values
summary(raster_1km_human_population_threshold_3)

#Step 2.2: Check the number of non-zero and NA values
num_zeros <- sum(values(raster_1km_human_population_threshold_3) == 0, na.rm = TRUE)
num_nas <- sum(is.na(values(raster_1km_human_population_threshold_3)))
num_non_zeros <- sum(values(raster_1km_human_population_threshold_3) > 0, na.rm = TRUE)
cat("Number of zero values:", num_zeros, "\n")
cat("Number of NA values:", num_nas, "\n")
cat("Number of non-zero values:", num_non_zeros, "\n")

#Step 2.3: Replace NA values with zeros (if appropriate for your analysis)
raster_1km_human_population_threshold_3[is.na(raster_1km_human_population_threshold_3)] <- 0
plot(raster_1km_human_population_threshold_3)
# Check if NA values are replaced
na_count_after <- sum(is.na(values(raster_1km_human_population_threshold_3)))
print(paste("Number of NA values after replacement:", na_count_after))

#Step 2.4: Get the number of rows and columns from the original raster
nrows <- nrow(raster_1km_human_population_threshold_3)
ncols <- ncol(raster_1km_human_population_threshold_3)

#Step 2.5: Convert the raster to a matrix and reshape it
population_matrix_LH <- matrix(values(raster_1km_human_population_threshold_3), 
                            nrow = nrows, ncol = ncols, byrow = TRUE)
# Check the dimensions and a small portion of the matrix
dim(population_matrix_LH)
print("Top-left corner of the matrix:")
print(population_matrix_LH[1:10, 1:10])  # Displaying the top-left 10x10 cells

# Step 3: Use the 'Fast-Fourier' transform to compute the convolution
convolved_matrix_LH <- convolve_image(population_matrix_LH, kernel_matrix_LH)

# Step 4: convert the matrix to raster

# Step 4.1: Convert the convolved matrix back to raster format using terra's rast function
convolved_raster_LH <- rast(nrows = nrow(raster_1km_human_population_threshold_3),
                                  ncols = ncol(raster_1km_human_population_threshold_3),
                                  xmin = xmin(raster_1km_human_population_threshold_3),
                                  xmax = xmax(raster_1km_human_population_threshold_3),
                                  ymin = ymin(raster_1km_human_population_threshold_3),
                                  ymax = ymax(raster_1km_human_population_threshold_3))

# Step 4.2: Assign the values of the convolved matrix back to the raster
values(convolved_raster_LH) <- convolved_matrix_LH

# Step 4.3: fix crs, extent and resolution
res(raster_1km_human_population_threshold_3)
res(convolved_raster_LH)
crs(raster_1km_human_population_threshold_3)
crs(convolved_raster_LH)
ext(raster_1km_human_population_threshold_3)
ext(convolved_raster_LH)

# Step 4.3: fix crs
crs(convolved_raster_LH) <- crs(raster_1km_human_population_threshold_3)

# Step 5: Plot results
plot(raster_1km_human_population_threshold_3, main = "Original Population Raster")
plot(kernel_matrix_LH, main = "Kernel Raster")
plot(convolved_raster_LH, main = "Convolved Raster with Identity Kernel")

# Step 6: mask the ocean

# Step 6.1: import the inland shapefile
inland_shapefile <- st_read("S:/Magali_Weissgerber/Europe frontiers_scratch/NUT3_from_rewilding_folder/Europe_limit_from_nut3_dissolve.shp")

# Step 6.2: mask the output raster
convolved_raster_masked_LH <- mask(convolved_raster_LH, inland_raster)
plot(convolved_raster_masked_LH)

# Compute quantile breaks (exclude NA values)
q_breaks <- quantile(values(convolved_raster_masked_LH), 
                     probs = seq(0, 1, length.out = 11), 
                     na.rm = TRUE)

# Plot with 10 quantile levels
plot(convolved_raster_masked_LH, 
     breaks = q_breaks, 
     col = hcl.colors(10, "YlOrRd", rev = TRUE), 
     main = "Convolved Raster (Quantile Breaks) large herbivores")

#Step 7: export
#output_file_WTT_masked_LH <- ("S:/Magali_Weissgerber/Species raster for social benefit of rewilding/WTT_LH_raster_masked_241212.tif")
#writeRaster(convolved_raster_masked_LH, output_file_WTT_masked_LH, overwrite=TRUE)

```

for each cell multiply the wildlife status (current presence where cell value = 1 and comeback probability where cell value is between 0 and 1) with the sum of people travelling

```{r}

#Occupancy raster (with 1, 0 and NA)
plot(raster_1km_wildlife_occupancy_stack)

#Probability based on the distance to 1 (cells with 1 have a 0 probability)
raster_1km_wildlife_probability_stack <- species_probability_stack_2
plot(raster_1km_wildlife_probability_stack)

#Number of humans travelling to see wildlife 
plot(convolved_raster_masked_LC)
plot(convolved_raster_masked_LH)

WTT_occ_proba_SB_stack <- c(convolved_raster_masked_LC, convolved_raster_masked_LH, raster_1km_wildlife_occupancy_stack, raster_1km_wildlife_probability_stack)
names(WTT_occ_proba_SB_stack)[1:2] <- c("WTT_LC", "WTT_LH")
plot(WTT_occ_proba_SB_stack)

#output_WTT_occ_proba_SB_stack <- ("S:/Magali_Weissgerber/Species raster for social benefit of rewilding/WTT_occ_proba_SB_stack_250912.tif")
#writeRaster(WTT_occ_proba_SB_stack, output_WTT_occ_proba_SB_stack, overwrite=TRUE)

```

Clean data for Social benefits map: number of people travelling, current mammals' distribution, potential future mammals' distribution

```{r}

# Check that the occurrence, probability and people travelling data fit the reference raster
compareGeom(WTT_occ_proba_SB_stack, ref_raster)
# Get the average number of social benefit (People travelling based on the Willingness-to-travel * probability of expansion). LC = large carnivores, LH = large herbivores
WTT_average = (WTT_occ_proba_SB_stack$WTT_LC + WTT_occ_proba_SB_stack$WTT_LH) / 2
plot(WTT_average) 

# Where species already as binary
# Current distribution binary (= with only 1s)
M_sum_binary <- ifel(WTT_occ_proba_SB_stack$M_sum > 0, 1, NA)
plot(M_sum_binary)

# Potential expansion with a combined probability of at least 0.5 (threshold)
P_SUM_threshold <- classify(WTT_occ_proba_SB_stack$P_SUM, matrix(c(-Inf, 0.5, 0, 0.5, Inf, 1), ncol = 3, byrow = TRUE))
# Plot to check
plot(P_SUM_threshold, main = "Binary probability (>0.5 = 1)")

```

## 3.3 - Determining social conflict

Import livestock distribution data

```{r}

# Get the total of cow, sheep, goat and pig per cell

# Import data

cow_population_path <- "I:/biocon/Gavin Stark/livestock and farmers maps/COWS/5_Ct_2015_Da.tif"
cow_population <- rast(cow_population_path)
plot(cow_population)

goat_population_path <- "I:/biocon/Gavin Stark/livestock and farmers maps/GOATS/5_Gt_2015_Da.tif"
goat_population <- rast(goat_population_path)
plot(goat_population)

sheep_population_path <- "I:/biocon/Gavin Stark/livestock and farmers maps/SHEEPS/5_Sh_2015_Da.tif"
sheep_population <- rast(sheep_population_path)
plot(sheep_population)

livestock_rasters <- c(cow_population, goat_population, sheep_population)

# Check extent and resolution, and crop to Europe

# Ensure CRS matches
livestock_rasters_aligned <- project(livestock_rasters, crs(inland_raster))

# Align extent and resolution of livestock rasters to inland_raster
livestock_rasters_aligned <- terra::resample(livestock_rasters_aligned, inland_raster, method = "bilinear")
plot(livestock_rasters_aligned)

# Mask the livestock rasters with the inland raster
livestock_rasters_masked <- mask(livestock_rasters_aligned, inland_raster)
plot(livestock_rasters_masked)
names(livestock_rasters_masked) <- c("livestock_Cattle", "livestock_Goat", "livestock_Sheep")

# It appear that is the row data Denmark is missing for goats
# Before summing the 3 layers the absence of Denmark for goats needs to be handled
# So we create a mask for valid land areas (non-NA areas in any raster)
land_mask <- !is.na(livestock_rasters_masked$livestock_Cattle) | !is.na(livestock_rasters_masked$livestock_Goat) | !is.na(livestock_rasters_masked$livestock_Sheep)

# We replace NA values with 0 so we can the sum the layer
conflict_raster_cow_sum <- livestock_rasters_masked$livestock_Cattle
conflict_raster_cow_sum[is.na(conflict_raster_cow_sum)] <- 0

conflict_raster_goat_sum <- livestock_rasters_masked$livestock_Goat
conflict_raster_goat_sum[is.na(conflict_raster_goat_sum)] <- 0

conflict_raster_sheep_sum <- livestock_rasters_masked$livestock_Sheep
conflict_raster_sheep_sum[is.na(conflict_raster_sheep_sum)] <- 0

```

An important livestock missing from the dataset are reindeers. However, there is not to my knowledge data on reindeer density in Europe, so I used the population estimations per country that I devided by the area reported for reindeers distribution.
Source: Moen, J., Forbes, B. C., Löf, A., & Horstkotte, T. (2022). Tipping points and regime shifts in reindeer husbandry: A systems approach. In Reindeer Husbandry and Global Environmental Change. Routledge.

```{r}

# 1. Load herding polygon and inland raster
herding_poly <- vect("S:/Magali_Weissgerber/Livestock for social conflict of rewilding/Reindeer/reindeer_polygon_250618.shp")
plot(herding_poly)
plot(inland_raster)
countries <- vect("I:/biocon/Projects/RewildingMAP/Consolidated/BASE_LAYERS/CountriesPanEURno500isl.shp")
plot(countries)

# 2. Reproject vector data to match inland_raster
herding_poly <- project(herding_poly, crs(inland_raster))
plot(herding_poly)
countries <- project(countries, crs(inland_raster))
names(countries)

# 3. Convert inland raster to polygon (to define valid inland zones)
inland_poly <- as.polygons(inland_raster, dissolve=TRUE, na.rm=TRUE)
plot(inland_poly)

# 4. Clip the herding polygon by inland extent
herding_poly <- st_as_sf(herding_poly)
inland_poly <- st_as_sf(inland_poly)
# Intersect
herding_clipped <- st_intersection(herding_poly, inland_poly)

plot(herding_clipped)

# 5. Filter countries to Norway, Sweden, Finland
nordics <- countries[countries$Country %in% c("Norway", "Sweden", "Finland"), ]
plot(nordics)
nordics <- st_as_sf(nordics)

# 6. Intersect clipped herding area with these three countries
herding_by_country <- st_intersection(herding_clipped, nordics)
plot(herding_by_country)
names(herding_by_country)

# 7. Recalculate area per polygon (in km²)
# Ensure projected CRS in meters (e.g., EPSG:3035)
herding_by_country <- st_transform(herding_by_country, crs = 3035)
herding_by_country$area_km2 <- as.numeric(st_area(herding_by_country)) / 1e6

# 8. Add reindeer population per country
herding_by_country$reindeer <- NA
herding_by_country$reindeer[herding_by_country$Country == "Sweden"] <- 241054
herding_by_country$reindeer[herding_by_country$Country == "Norway"] <- 210000
herding_by_country$reindeer[herding_by_country$Country == "Finland"] <- 190000

# 9. Calculate reindeer density per km²
herding_by_country$density <- herding_by_country$reindeer / herding_by_country$area_km2

# 10. Create raster template (same resolution/extent as inland_raster)
template <- inland_raster

# Convert to SpatVector for terra::rasterize
herding_vect <- vect(herding_by_country)

# 11. Rasterize using the density field
density_raster <- rasterize(herding_vect, template, field = "density", background = 0)
names(density_raster) <- "livestock_reindeer"
plot(density_raster)
reindeer_population <- density_raster

# Mask it to land
reindeer_population_masked <- mask(density_raster, land_mask, maskvalues = FALSE)
plot(reindeer_population_masked)

```

Prepare one layer with all livestock species

```{r}

# Stack the rasters
conflict_raster_stack_summed <- sum(conflict_raster_cow_sum, conflict_raster_goat_sum, conflict_raster_sheep_sum, reindeer_population)

# Apply the land mask to retain original NA areas
conflict_raster_stack_summed <- mask(conflict_raster_stack_summed, land_mask, maskvalues=FALSE)

# Have a look
print(conflict_raster_stack_summed)

# Stack all the rasters
conflict_raster_stack_complete <- c(livestock_rasters_masked, reindeer_population_masked, conflict_raster_stack_summed)
names(conflict_raster_stack_complete) <- c("livestock_cow", "livestock_goat", "livestock_sheep", "livestock_reindeer", "livestock_sum")
plot(conflict_raster_stack_complete)

```

To visualize the conflicts we consider livestock and carnivores Canis lupus and Ursus arctos.
Here are clean data for social conflicts: 1) livestock density, 2) current Canis lupus and Ursus arctos distribution, 3) potential Canis lupus and Ursus arctos distribution

```{r}

# livestock population as continuous
plot(conflict_raster_stack_complete$livestock_sum)

# requires species (bear and wolf) expansion as binary (threshold at 0.5) and livestock population as continuous

# Sum CANLUP and URSARC proba
P_CN_UA <- WTT_occ_proba_SB_stack$P_CANLUP + WTT_occ_proba_SB_stack$P_URSARC
plot(P_CN_UA)
# Potential expansion with a combined probability of at least 0.5 (threshold)
P_CN_UA_threshold <- classify(P_CN_UA, matrix(c(-Inf, 0.5, 0, 0.5, Inf, 1), ncol = 3, byrow = TRUE))
# Plot to check
plot(P_CN_UA_threshold, main = "Binary probability (≥0.5 = 1)")

# Where species (bear and wolf) already there as binary
# Sum CANLUP and URSARC current
M_CN_UA <- WTT_occ_proba_SB_stack$M_CANLUP + WTT_occ_proba_SB_stack$M_URSARC
plot(M_CN_UA)
# Current distribution binary (= with only 1s)
M_CN_UA_binary <- ifel(M_CN_UA > 0, 1, NA)
plot(M_CN_UA_binary)

```

## 3.4: Export and import each layer so there is no need to process all the data each time

```{r}

output_dir = "S:/Magali_Weissgerber/Data_framework_verifying_260324"

# Climate change mitigation
  
writeRaster(hanpp_masked, filename = file.path(output_dir, "layers_maps_hanpp_masked_260324.tif"), overwrite = TRUE)
hanpp_masked <- rast(file.path(output_dir, "layers_maps_hanpp_masked_260324.tif"))

writeRaster(abandonment_masked, filename = file.path(output_dir, "layers_maps_abandonment_masked_260324.tif"), overwrite = TRUE)
abandonment_masked <- rast(file.path(output_dir, "layers_maps_abandonment_masked_260324.tif"))

writeRaster(pot_carbon_masked, filename = file.path(output_dir, "layers_maps_pot_carbon_masked_260324.tif"), overwrite = TRUE)
pot_carbon_masked <- rast(file.path(output_dir, "layers_maps_pot_carbon_masked_260324.tif"))

# Climate change adaptation

writeRaster(high_meff_masked, filename = file.path(output_dir, "layers_maps_high_meff_masked_260324.tif"), overwrite = TRUE)
high_meff_masked <- rast(file.path(output_dir, "layers_maps_high_meff_masked_260324.tif"))

writeRaster(cv_masked, filename = file.path(output_dir, "layers_maps_cv_masked_260324.tif"), overwrite = TRUE)
cv_masked <- rast(file.path(output_dir, "layers_maps_cv_masked_260324.tif"))

# Social benefits

writeRaster(WTT_average, filename = file.path(output_dir, "layers_maps_WTT_average_260324.tif"), overwrite = TRUE)
WTT_average <- rast(file.path(output_dir, "layers_maps_WTT_average_260324.tif"))

writeRaster(M_sum_binary, filename = file.path(output_dir, "layers_maps_M_sum_binary_260324.tif"), overwrite = TRUE)
M_sum_binary <- rast(file.path(output_dir, "layers_maps_M_sum_binary_260324.tif"))

writeRaster(P_SUM_threshold, filename = file.path(output_dir, "layers_maps_P_SUM_threshold_260324.tif"), overwrite = TRUE)
P_SUM_threshold <- rast(file.path(output_dir, "layers_maps_P_SUM_threshold_260324.tif"))

# Social Conflict

livestock_sum <- conflict_raster_stack_complete$livestock_sum
writeRaster(livestock_sum, filename = file.path(output_dir, "layers_maps_livestock_sum_260324.tif"), overwrite = TRUE)
livestock_sum <- rast(file.path(output_dir, "layers_maps_livestock_sum_260324.tif"))

writeRaster(M_CN_UA_binary, filename = file.path(output_dir, "layers_maps_M_CN_UA_binary_260324.tif"), overwrite = TRUE)
M_CN_UA_binary <- rast(file.path(output_dir, "layers_maps_M_CN_UA_binary_260324.tif"))

writeRaster(P_CN_UA_threshold, filename = file.path(output_dir, "layers_maps_P_CN_UA_threshold_260324.tif"), overwrite = TRUE)
P_CN_UA_threshold <- rast(file.path(output_dir, "layers_maps_P_CN_UA_threshold_260324.tif"))

```

## II. map each data individually (to prepare the figure in supplementary materials)

# 1. Current

Low land use intensity = natural disturbance regime

```{r}

plot(hanpp_masked)

# Convert raster to data frame for ggplot
df_hanpp <- as.data.frame(hanpp_masked, xy = TRUE)
names(df_hanpp)[3] <- "value"

# If it's binary (0/1), convert to factor
df_hanpp$value <- factor(df_hanpp$value, levels = c(0, 1))

# Define colors
nat_colors <- c(
  "0" = "#FFFFFF",  # white for 0
  "1" = "#f29842"   # dark green for natural areas
)

nat_labels <- c(
  "0" = "Other land",
  "1" = "Natural disturbance regime"
)

# Plot with ggplot
map_hanpp <- ggplot(df_hanpp, aes(x = x, y = y, fill = value)) +
  geom_tile() +
  scale_fill_manual(
    values = nat_colors,
    breaks = c("1"),
    labels = nat_labels["1"],   # only show "Natural areas" in legend
    name = "Naturalness"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "High naturalness in Europe")

ggsave(
  file.path(output_dir, 'intermediary_map_intensity_260324.png'),
  plot = map_hanpp,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

Effective mesh size

```{r}

plot(high_meff_masked)

# Convert raster to data frame for ggplot
df_meff <- as.data.frame(high_meff_masked, xy = TRUE)
names(df_meff)[3] <- "value"

# If it's binary (0/1), convert to factor
df_meff$value <- factor(df_meff$value, levels = c(0, 1))

# Define colors
meff_colors <- c(
  "0" = "#FFFFFF",  # white for 0
  "1" = "#f29842"   # orange
)

meff_labels <- c(
  "0" = "Other land",
  "1" = "High meff"
)

# Plot with ggplot
map_meff <- ggplot(df_meff, aes(x = x, y = y, fill = value)) +
  geom_tile() +
  scale_fill_manual(
    values = meff_colors,
    breaks = c("1"),
    labels = meff_labels["1"],   # only show orange in legend
    name = "Effective mesh size"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "High effective mesh size in Europe")

ggsave(
  file.path(output_dir, 'intermediary_map_meff_260324.png'),
  plot = map_meff,
  width = 26, height = 25, units = 'cm', dpi = 300
)


```

Current 8 mammal species for social benefit

```{r}

plot(M_sum_binary)

# Convert raster to data frame for ggplot
df_M_sum_binary <- as.data.frame(M_sum_binary, xy = TRUE)
names(df_M_sum_binary)[3] <- "value"

# If it's binary (0/1), convert to factor
df_M_sum_binary$value <- factor(df_M_sum_binary$value, levels = c(0, 1))

# Define colors
ecotourism_colors <- c(
  "0" = "#FFFFFF",  # white for 0
  "1" = "#f29842"   # orange
)

ecotourism_labels <- c(
  "0" = "Other land",
  "1" = "8 large mammal species"
)

# Plot with ggplot
map_species_current <- ggplot(df_M_sum_binary, aes(x = x, y = y, fill = value)) +
  geom_tile() +
  scale_fill_manual(
    values = ecotourism_colors,
    breaks = c("1"),
    labels = ecotourism_labels["1"],   # only show orange in legend
    name = "Current distribution without expansion"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "Large mammals current distribution without expansion in Europe")

ggsave(
  file.path(output_dir, 'intermediary_map_M_sum_binary_260324.png'),
  plot = map_species_current,
  width = 26, height = 25, units = 'cm', dpi = 300
)


```

Carnivores current

```{r}

plot(M_CN_UA_binary)

# Convert raster to data frame for ggplot
df_M_CN_UA_binary <- as.data.frame(M_CN_UA_binary, xy = TRUE)
names(df_M_CN_UA_binary)[3] <- "value"

# If it's binary (0/1), convert to factor
df_M_CN_UA_binary$value <- factor(df_M_CN_UA_binary$value, levels = c(0, 1))

# Define colors
carnivores_colors <- c(
  "0" = "#FFFFFF",  # white for 0
  "1" = "#f29842"   # orange
)

carnivores_labels <- c(
  "0" = "Other land",
  "1" = "Canis lupus lupus & Ursus arctos"
)

# Plot with ggplot
map_carnivores <- ggplot(df_M_CN_UA_binary, aes(x = x, y = y, fill = value)) +
  geom_tile() +
  scale_fill_manual(
    values = carnivores_colors,
    breaks = c("1"),
    labels = carnivores_labels["1"],   # only show orange in legend
    name = "Current distribution without expansion"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "Large carnivores current distribution without expansion in Europe")

ggsave(
  file.path(output_dir, 'intermediary_map_current_carnivores_260324.png'),
  plot = map_carnivores,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

# 2. Potential

Abandonment

```{r}

plot(abandonment_masked)

# Convert raster to data frame for ggplot
df_aband <- as.data.frame(abandonment_masked, xy = TRUE)
names(df_aband)[3] <- "value"

# If it's binary (0/1), convert to factor
df_aband$value <- factor(df_aband$value, levels = c(0, 1))

# Define colors
aband_colors <- c(
  "0" = "#FFFFFF",  # white for 0
  "1" = "#AB47BC"   # purple
)

aband_labels <- c(
  "0" = "Other land",
  "1" = "Potential abandonment"
)

# Plot with ggplot
map_aband <- ggplot(df_aband, aes(x = x, y = y, fill = value)) +
  geom_tile() +
  scale_fill_manual(
    values = aband_colors,
    breaks = c("1"),
    labels = aband_labels["1"],   # only show purple in legend
    name = "Abandonment status"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "Areas potentially abandonned")

ggsave(
  file.path(output_dir, 'intermediary_map_aband_260324.png'),
  plot = map_aband,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

Species probable expansion

```{r}

plot(P_SUM_threshold)

# Convert raster to data frame for ggplot
df_mam_exp <- as.data.frame(P_SUM_threshold, xy = TRUE)
names(df_mam_exp)[3] <- "value"

# If it's binary (0/1), convert to factor
df_mam_exp$value <- factor(df_mam_exp$value, levels = c(0, 1))

# Define colors
mam_exp_colors <- c(
  "0" = "#FFFFFF",  # white for 0
  "1" = "#AB47BC"   # purple
)

mam_exp_labels <- c(
  "0" = "Other land",
  "1" = "Proba >= 0.5"
)

# Plot with ggplot
map_mam_exp <- ggplot(df_mam_exp, aes(x = x, y = y, fill = value)) +
  geom_tile() +
  scale_fill_manual(
    values = mam_exp_colors,
    breaks = c("1"),
    labels = mam_exp_labels["1"],   # only show purple in legend
    name = "Potential mammal expansion"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "Potential expansion of 8 mammal species")

ggsave(
  file.path(output_dir, 'intermediary_map_mam_exp_260324.png'),
  plot = map_mam_exp,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

Carnivores probable expansion

```{r}

plot(P_CN_UA_threshold)

# Convert raster to data frame for ggplot
df_carn_exp <- as.data.frame(P_CN_UA_threshold, xy = TRUE)
names(df_carn_exp)[3] <- "value"

# If it's binary (0/1), convert to factor
df_carn_exp$value <- factor(df_carn_exp$value, levels = c(0, 1))

# Define colors
carn_exp_colors <- c(
  "0" = "#FFFFFF",  # white for 0
  "1" = "#AB47BC"   # purple
)

carn_exp_labels <- c(
  "0" = "Other land",
  "1" = "Proba >= 0.5"
)

# Plot with ggplot
map_carn_exp <- ggplot(df_carn_exp, aes(x = x, y = y, fill = value)) +
  geom_tile() +
  scale_fill_manual(
    values = carn_exp_colors,
    breaks = c("1"),
    labels = carn_exp_labels["1"],   # only show purple in legend
    name = "Potential carnivore expansion"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "Potential expansion of Canis lupus lupus and Ursus arctos")

ggsave(
  file.path(output_dir, 'intermediary_map_carn_exp_260324.png'),
  plot = map_carn_exp,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

# 3. Benefits / proxy of outcome (quantiles)

Carbon potential

```{r}

plot(pot_carbon_masked)

# Step 1: Compute quantile breaks (10 bins)
breaks_carbon <- quantile(values(pot_carbon_masked), probs = seq(0, 1, length.out = 11), na.rm = TRUE)

# Step 2: Classify into quantile bins (1–10)
carbon_class <- classify(
  pot_carbon_masked,
  rcl = matrix(c(
    breaks_carbon[1],  breaks_carbon[2],  1,
    breaks_carbon[2],  breaks_carbon[3],  2,
    breaks_carbon[3],  breaks_carbon[4],  3,
    breaks_carbon[4],  breaks_carbon[5],  4,
    breaks_carbon[5],  breaks_carbon[6],  5,
    breaks_carbon[6],  breaks_carbon[7],  6,
    breaks_carbon[7],  breaks_carbon[8],  7,
    breaks_carbon[8],  breaks_carbon[9],  8,
    breaks_carbon[9],  breaks_carbon[10], 9,
    breaks_carbon[10], breaks_carbon[11], 10
  ), ncol = 3, byrow = TRUE)
)

# Step 3: Convert to data frame
df_carbon <- as.data.frame(carbon_class, xy = TRUE)
names(df_carbon)[3] <- "bin"

# Step 4: Map with grey gradient
map_carbon <- ggplot(df_carbon, aes(x = x, y = y, fill = bin)) +
  geom_tile() +
  scale_fill_gradientn(
    colors = gray.colors(10, start = 0.9, end = 0.2), # light grey (0.9) → dark grey (0.2)
    breaks = 1:10,
    labels = round(breaks_carbon[-1], 2),   # numeric labels from quantile breaks
    name = "Potential carbon storage (tC/km2)"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "Potential carbon storage (10-quantile map)")

ggsave(
  file.path(output_dir, 'intermediary_map_carbon_pot_260324.png'),
  plot = map_carbon,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

Climate velocity

```{r}

plot(cv_masked)

# Step 1: Compute quantile breaks (10 bins)
breaks_cv <- quantile(values(cv_masked), probs = seq(0, 1, length.out = 11), na.rm = TRUE)

# Step 2: Classify into quantile bins (1–10)
cv_class <- classify(
  cv_masked,
  rcl = matrix(c(
    breaks_cv[1],  breaks_cv[2],  1,
    breaks_cv[2],  breaks_cv[3],  2,
    breaks_cv[3],  breaks_cv[4],  3,
    breaks_cv[4],  breaks_cv[5],  4,
    breaks_cv[5],  breaks_cv[6],  5,
    breaks_cv[6],  breaks_cv[7],  6,
    breaks_cv[7],  breaks_cv[8],  7,
    breaks_cv[8],  breaks_cv[9],  8,
    breaks_cv[9],  breaks_cv[10], 9,
    breaks_cv[10], breaks_cv[11], 10
  ), ncol = 3, byrow = TRUE)
)

# Step 3: Convert to data frame
df_cv <- as.data.frame(cv_class, xy = TRUE)
names(df_cv)[3] <- "bin"

# Step 4: Map with grey gradient
map_cv <- ggplot(df_cv, aes(x = x, y = y, fill = bin)) +
  geom_tile() +
  scale_fill_gradientn(
    colors = gray.colors(10, start = 0.9, end = 0.2), # light grey (0.9) → dark grey (0.2)
    breaks = 1:10,
    labels = round(breaks_cv[-1], 2),   # numeric labels from quantile breaks
    name = "Climate velocity (km/year)"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "Climate velocity (10-quantile map)")

ggsave(
  file.path(output_dir, 'intermediary_map_cv_260324.png'),
  plot = map_cv,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

Number of people travelling

```{r}

plot(WTT_average)

# Step 1: Compute quantile breaks (10 bins)
breaks_WTT <- quantile(values(WTT_average), probs = seq(0, 1, length.out = 11), na.rm = TRUE)

# Step 2: Classify into quantile bins (1–10)
WTT_class <- classify(
  WTT_average,
  rcl = matrix(c(
    breaks_WTT[1],  breaks_WTT[2],  1,
    breaks_WTT[2],  breaks_WTT[3],  2,
    breaks_WTT[3],  breaks_WTT[4],  3,
    breaks_WTT[4],  breaks_WTT[5],  4,
    breaks_WTT[5],  breaks_WTT[6],  5,
    breaks_WTT[6],  breaks_WTT[7],  6,
    breaks_WTT[7],  breaks_WTT[8],  7,
    breaks_WTT[8],  breaks_WTT[9],  8,
    breaks_WTT[9],  breaks_WTT[10], 9,
    breaks_WTT[10], breaks_WTT[11], 10
  ), ncol = 3, byrow = TRUE)
)

# Step 3: Convert to data frame
df_WTT <- as.data.frame(WTT_class, xy = TRUE)
names(df_WTT)[3] <- "bin"

# Step 4: Map with grey gradient
map_WTT <- ggplot(df_WTT, aes(x = x, y = y, fill = bin)) +
  geom_tile() +
  scale_fill_gradientn(
    colors = gray.colors(10, start = 0.9, end = 0.2), # light grey (0.9) → dark grey (0.2)
    breaks = 1:10,
    labels = round(breaks_WTT[-1], 2),   # numeric labels from quantile breaks
    name = "People travelling (count/km2)"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "Number of people willing to travel (10-quantile map)")

ggsave(
  file.path(output_dir, 'intermediary_map_WTT_260304.png'),
  plot = map_WTT,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

Livestock

```{r}

plot(livestock_sum)

# Step 1: Compute quantile breaks (10 bins)
breaks_livestock <- quantile(values(livestock_sum), probs = seq(0, 1, length.out = 11), na.rm = TRUE)

# Step 2: Classify into quantile bins (1–10)
livestock_class <- classify(
  livestock_sum,
  rcl = matrix(c(
    breaks_livestock[1],  breaks_livestock[2],  1,
    breaks_livestock[2],  breaks_livestock[3],  2,
    breaks_livestock[3],  breaks_livestock[4],  3,
    breaks_livestock[4],  breaks_livestock[5],  4,
    breaks_livestock[5],  breaks_livestock[6],  5,
    breaks_livestock[6],  breaks_livestock[7],  6,
    breaks_livestock[7],  breaks_livestock[8],  7,
    breaks_livestock[8],  breaks_livestock[9],  8,
    breaks_livestock[9],  breaks_livestock[10], 9,
    breaks_livestock[10], breaks_livestock[11], 10
  ), ncol = 3, byrow = TRUE)
)

# Step 3: Convert to data frame
df_livestock <- as.data.frame(livestock_class, xy = TRUE)
names(df_livestock)[3] <- "bin"

# Step 4: Map with inverted grey gradient
map_livestock <- ggplot(df_livestock, aes(x = x, y = y, fill = bin)) +
  geom_tile() +
  scale_fill_gradientn(
    colors = gray.colors(10, start = 0.2, end = 0.9), # dark grey (low) → light grey (high)
    breaks = 1:10,
    labels = round(breaks_livestock[-1], 2),   # numeric labels from quantile breaks
    name = "Livestock density (count/km²)"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "Livestock density (sheep, cow, goat, reindeer) (10-quantile map)")

ggsave(
  file.path(output_dir, 'intermediary_map_livestock_260304.png'),
  plot = map_livestock,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

## III. produce the bivariate maps

Climate change mitigation (potential carbon x (abandonment and low land use intensity))

```{r}

# Step 1: Round NPP and classify into 10 quantiles
pot_carbon_ceil <- ceiling(pot_carbon_masked)
pot_carbon_breaks <- quantile(values(pot_carbon_ceil), probs = seq(0, 1, length.out = 11), na.rm = TRUE)

pot_carbon_classified <- classify(pot_carbon_ceil, rcl = matrix(c(
  pot_carbon_breaks[1] - 0.00001, pot_carbon_breaks[2], 1,
  pot_carbon_breaks[2], pot_carbon_breaks[3], 2,
  pot_carbon_breaks[3], pot_carbon_breaks[4], 3,
  pot_carbon_breaks[4], pot_carbon_breaks[5], 4,
  pot_carbon_breaks[5], pot_carbon_breaks[6], 5,
  pot_carbon_breaks[6], pot_carbon_breaks[7], 6,
  pot_carbon_breaks[7], pot_carbon_breaks[8], 7,
  pot_carbon_breaks[8], pot_carbon_breaks[9], 8,
  pot_carbon_breaks[9], pot_carbon_breaks[10], 9,
  pot_carbon_breaks[10], pot_carbon_breaks[11], 10
), ncol = 3, byrow = TRUE))

# Step 2: Mutually exclusive masks
abandonment_mask <- ifel(abandonment_masked == 1, 1, 0)
plot(abandonment_mask)
hanpp_mask <- ifel((hanpp_masked == 1) & (abandonment_masked != 1), 1, 0)
plot(hanpp_mask)
mitigation_residual_mask <- ifel((abandonment_masked != 1) & (hanpp_masked != 1), 1, 0)
plot(mitigation_residual_mask)

# Step 3: Offset and combine values by category
abandonment_layer_mitigation <- ifel(abandonment_mask == 1, pot_carbon_classified + 10, NA)  # 11–20
plot(abandonment_layer_mitigation)
hanpp_layer <- ifel(hanpp_mask == 1, pot_carbon_classified + 20, NA)              # 21–30
plot(hanpp_layer)
no_abandonment_layer <- ifel(mitigation_residual_mask == 1, pot_carbon_classified, NA)       # 1–10
plot(no_abandonment_layer)

# Step 4: Merge into a single raster
pot_carbon_abandonment_combined <- merge(abandonment_layer_mitigation, hanpp_layer, no_abandonment_layer)
par(mfrow = c(1,1))
plot(pot_carbon_abandonment_combined)

# Step 5: Convert to dataframe
pot_carbon_abandonment_df <- as.data.frame(pot_carbon_abandonment_combined, xy = TRUE)
colnames(pot_carbon_abandonment_df) <- c("x", "y", "value")

# Step 6: Custom 30-color palette
abandon_palette <- c(
  # 1–10: Greys for no abandonment
  "#FAFAFA", "#F0F0F0", "#E0E0E0", "#D6D6D6", "#C4C4C4",
  "#B0B0B0", "#9E9E9E", "#7D7D7D", "#616161", "#424242",
  
  # 11–20: Purple for abandonment
  "#F3E5F5", "#E1BEE7", "#CE93D8", "#BA68C8", "#AB47BC",
  "#9C27B0", "#8E24AA", "#7B1FA2", "#6A1B9A", "#4A148C",
  
  # 21–30: Browns for HANPP
  "#fff4e3", "#fee0b6", "#fdd697", "#fdb863", "#f29842",
  "#e08214", "#cc6d0d", "#b35806", "#a34906", "#7f3b08"
)

# Step 7: Map value to colors
pot_carbon_abandonment_df <- pot_carbon_abandonment_df %>%
  mutate(color = ifelse(!is.na(value), abandon_palette[value], "#d3d3d3"))

# Step 8: Map
pot_carbon_abandonment_map <- ggplot() +
  geom_tile(data = pot_carbon_abandonment_df, aes(x = x, y = y, fill = color)) +
  scale_fill_identity() +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  ) +
  labs(
    title = "Climate Change Mitigation on Abandoned Farmlands"
  ) +
  theme(
    plot.title = element_text(
      size = 18,        # font size (increase/decrease as needed)
      face = "bold",    # makes it bold
      hjust = 0.5       # centers the title
    )
  )

# Step 9: Save map
# png
ggsave(
  file.path(output_dir, 'pot_carbon_abandonment_hanpp_map_260324.png'),
  plot = pot_carbon_abandonment_map,
  width = 26, height = 25, units = 'cm', dpi = 300
)

# LEGEND SECTION ------------------------------------------------------------

# Step 10: Legend
legend_df_pot_carbon <- expand.grid(
  base_level = 1:10,
  category = c("No abandonment", "Abandonment", "HANPP")
)

legend_df_pot_carbon <- legend_df_pot_carbon %>%
  mutate(
    category = recode(
      category,
      "No abandonment" = "No future abandonment",
      "Abandonment" = "Probable future abandonment",
      "HANPP" = "Low land use intensity"
    ),
    color = case_when(
      category == "No future abandonment" ~ abandon_palette[base_level],
      category == "Probable future abandonment" ~ abandon_palette[base_level + 10],
      category == "Low land use intensity" ~ abandon_palette[base_level + 20]
    )
  )

pot_carbon_legend_limits <- round(pot_carbon_breaks)

pot_carbon_abandonment_legend <- ggplot(legend_df_pot_carbon, aes(x = base_level, y = category, fill = color)) +
  geom_tile() +
  scale_fill_identity() +
  scale_x_continuous(
    breaks = seq(0.5, 10.5, by = 1),
    labels = pot_carbon_legend_limits,
    position = 'top'
  ) +
  labs(
    x = "Potential carbon storage (tC/km2)",
    y = "Land status"
  ) +
  theme_minimal() +
  theme(
    axis.title.y = element_text(size = 14, angle = 0, vjust = 1.6, hjust = 1, margin = margin(r = -80)),
    axis.text = element_text(size = 11),
    axis.text.x = element_text(size = 11, angle = 90),
    axis.text.x.top = element_text(size = 11, hjust = 0),
    axis.title.x.top = element_text(size = 14, margin = margin(b = 20)),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(10, 10, 10, 100)
  ) +
  coord_fixed(ratio = 0.5)

plot(pot_carbon_abandonment_legend)

ggsave(
  file.path(output_dir, 'pot_carbon_abandonment_hanpp_legend_260324.png'),
  plot = pot_carbon_abandonment_legend,
  width = 24, height = 10, units = 'cm', dpi = 300
)

# The figure in the article is built upon the maps and legend as separate .png files, here are two supplementary steps to have both together

############ Combine map and legend

# Combine the fixed legend with the map
final_plot_pot_carbon <- cowplot::plot_grid(
  pot_carbon_abandonment_map,
  pot_carbon_abandonment_legend,
  ncol = 2
)

# Save final composite
ggsave(
  file.path(output_dir, "pot_carbon_abandonment_hanpp_map_with_legend_260324.png"),
  plot = final_plot_pot_carbon,
  width = 40, height = 25, units = "cm", dpi = 300
)


```

Climate change adaptation (climate velocity x (abandonment and effective mesh size))

```{r}

# Step 1: Round CV and classify into 10 quantiles
cv_ceil <- ceiling(cv_masked)
plot(cv_masked)
cv_breaks <- quantile(values(cv_ceil), probs = seq(0, 1, length.out = 11), na.rm = TRUE)

cv_classified <- classify(cv_ceil, rcl = matrix(c(
  cv_breaks[1] - 0.00001, cv_breaks[2], 1,
  cv_breaks[2], cv_breaks[3], 2,
  cv_breaks[3], cv_breaks[4], 3,
  cv_breaks[4], cv_breaks[5], 4,
  cv_breaks[5], cv_breaks[6], 5,
  cv_breaks[6], cv_breaks[7], 6,
  cv_breaks[7], cv_breaks[8], 7,
  cv_breaks[8], cv_breaks[9], 8,
  cv_breaks[9], cv_breaks[10], 9,
  cv_breaks[10], cv_breaks[11], 10
), ncol = 3, byrow = TRUE))

# Step 2: Mutually exclusive masks
abandonment_mask <- ifel(abandonment_masked == 1, 1, 0)
meff_mask <- ifel((high_meff_masked == 1) & (abandonment_masked != 1), 1, 0)
cv_residual_mask <- ifel((abandonment_masked != 1) & (high_meff_masked != 1), 1, 0)

# Step 3: Offset and combine values by category
abandonment_layer_adaptation <- ifel(abandonment_mask == 1, cv_classified + 10, NA)  # 11–20
meff_layer <- ifel(meff_mask == 1, cv_classified + 20, NA)                # 21–30
no_abandonment_layer <- ifel(cv_residual_mask == 1, cv_classified, NA)       # 1–10

# Step 4: Merge into a single raster
cv_abandonment_combined <- merge(abandonment_layer_adaptation, meff_layer, no_abandonment_layer)
plot(cv_abandonment_combined)

# Step 5: Convert to dataframe
cv_abandonment_df <- as.data.frame(cv_abandonment_combined, xy = TRUE)
colnames(cv_abandonment_df) <- c("x", "y", "value")

# Step 6: Custom 30-color palette
abandon_palette <- c(
  # 1–10: Greys for no abandonment
  "#FAFAFA", "#F0F0F0", "#E0E0E0", "#D6D6D6", "#C4C4C4",
  "#B0B0B0", "#9E9E9E", "#7D7D7D", "#616161", "#424242",
  # 11–20: Purple for abandonment
  "#F3E5F5", "#E1BEE7", "#CE93D8", "#BA68C8", "#AB47BC",
  "#9C27B0", "#8E24AA", "#7B1FA2", "#6A1B9A", "#4A148C",
  # 21–30: Browns for MEFF
  "#fff4e3", "#fee0b6", "#fdd697", "#fdb863", "#f29842",
  "#e08214", "#cc6d0d", "#b35806", "#a34906", "#7f3b08"
)

# Step 7: Map value to colors
cv_abandonment_df <- cv_abandonment_df %>%
  mutate(color = ifelse(!is.na(value), abandon_palette[value], "#d3d3d3"))

# Step 8: Map
cv_abandonment_map <- ggplot() +
  geom_tile(data = cv_abandonment_df, aes(x = x, y = y, fill = color)) +
  scale_fill_identity() +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )  +
  labs(
    title = "Climate Change Adaptation for Ecosystem Resilience"
  ) +
  theme(
    plot.title = element_text(
      size = 18,        # font size (increase/decrease as needed)
      face = "bold",    # makes it bold
      hjust = 0.5       # centers the title
    ))
  

# Step 9: Save map
ggsave(
  file.path(output_dir, 'cv_abandonment_meff_map_260324.png'),
  plot = cv_abandonment_map,
  width = 26, height = 25, units = 'cm', dpi = 300
)

# LEGEND SECTION ------------------------------------------------------------

# Step 10: Legend
legend_df_cv <- expand.grid(
  base_level = 1:10,
  category = c("No abandonment", "Abandonment", "MEFF")
)

legend_df_cv <- legend_df_cv %>%
  mutate(
    category = recode(
      category,
      "No abandonment" = "No future abandonment",
      "Abandonment" = "Probable future abandonment",
      "MEFF" = "Current high connectivity"
    ),
    color = case_when(
      category == "No future abandonment" ~ abandon_palette[base_level],
      category == "Probable future abandonment" ~ abandon_palette[base_level + 10],
      category == "Current high connectivity" ~ abandon_palette[base_level + 20]
    )
  )

cv_legend_limits <- round(cv_breaks)

cv_abandonment_legend <- ggplot(legend_df_cv, aes(x = base_level, y = category, fill = color)) +
  geom_tile() +
  scale_fill_identity() +
  scale_x_continuous(
    breaks = seq(0.5, 10.5, by = 1),
    labels = cv_legend_limits,
    position = 'top'
  ) +
  labs(
    x = "Climate velocity (km/year)",
    y = "Land status"
  ) +
  theme_minimal() +
  theme(
    axis.title.y = element_text(size = 14, angle = 0, vjust = 1.6, hjust = 1, margin = margin(r = -80)),
    axis.text = element_text(size = 11),
    axis.text.x = element_text(size = 11, angle = 90),
    axis.text.x.top = element_text(size = 11, hjust = 0),
    axis.title.x.top = element_text(size = 14, margin = margin(b = 20)),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(10, 10, 10, 100)
  ) +
  coord_fixed(ratio = 0.5)

plot(cv_abandonment_legend)

ggsave(
  file.path(output_dir, 'cv_abandonment_meff_legend_260324.png'),
  plot = cv_abandonment_legend,
  width = 24, height = 10, units = 'cm', dpi = 300
)

# The figure in the article is built upon the maps and legend as separate .png files, here are two supplementary steps to have both together

############ Combine map and legend

# Combine the fixed legend with the map
final_plot_cv <- cowplot::plot_grid(
  cv_abandonment_map,
  cv_abandonment_legend,
  ncol = 2
)

# Save final composite
ggsave(
  file.path(output_dir, "cv_abandonment_meff_map_with_legend_260324.png"),
  plot = final_plot_cv,
  width = 40, height = 25, units = "cm", dpi = 300
)


```

Social benefits (people travelling x (species current distribution and potential distribution))

```{r}

# Bivariate mapping: binary species expansion x 10 levels of people travelling

# Step 0: remove overlap between current and proba
M_sum_binary_sub <- M_sum_binary - P_SUM_threshold
M_sum_binary_sub <- ifel(M_sum_binary_sub == -1, 0, M_sum_binary_sub)
plot(M_sum_binary_sub)

# Step 1: Round NPP and classify into 10 quantiles
WTT_ceil <- ceiling(WTT_average)
WTT_breaks <- quantile(values(WTT_ceil), probs = seq(0, 1, length.out = 11), na.rm = TRUE)

WTT_classified <- classify(WTT_ceil, rcl = matrix(c(
  WTT_breaks[1] - 0.00001, WTT_breaks[2], 1,
  WTT_breaks[2], WTT_breaks[3], 2,
  WTT_breaks[3], WTT_breaks[4], 3,
  WTT_breaks[4], WTT_breaks[5], 4,
  WTT_breaks[5], WTT_breaks[6], 5,
  WTT_breaks[6], WTT_breaks[7], 6,
  WTT_breaks[7], WTT_breaks[8], 7,
  WTT_breaks[8], WTT_breaks[9], 8,
  WTT_breaks[9], WTT_breaks[10], 9,
  WTT_breaks[10], WTT_breaks[11], 10
), ncol = 3, byrow = TRUE))

# Step 2: Combine species logic for binary mapping (exclusive values)
# Prioritize M_SUM > P_SUM > No species
WTT_M_sum <- ifel(M_sum_binary_sub == 1 & P_SUM_threshold == 0, WTT_classified + 20, NA)
WTT_P_sum <- ifel(P_SUM_threshold == 1 & M_sum_binary_sub == 0, WTT_classified + 10, NA)
WTT_no_species <- ifel(M_sum_binary_sub == 0 & P_SUM_threshold == 0, WTT_classified, NA)

# Step 3: Combine using cover() to prioritize non-NA in order
WTT_species_combined <- merge(WTT_no_species, WTT_P_sum, WTT_M_sum)
plot(WTT_species_combined)

# Step 4: Convert to dataframe
WTT_species_df <- as.data.frame(WTT_species_combined, xy = TRUE)
colnames(WTT_species_df) <- c("x", "y", "value")

# Step 5: Define 30-class palette
abandon_palette <- c(
  # 1-10: Shades of grey for no abandonment
  "#FAFAFA", "#F0F0F0", "#E0E0E0", "#D6D6D6", "#C4C4C4",
  "#B0B0B0", "#9E9E9E", "#7D7D7D", "#616161", "#424242",
  # 11-20: Shades of purple for P_SUM species
  "#F3E5F5", "#E1BEE7", "#CE93D8", "#BA68C8", "#AB47BC",
  "#9C27B0", "#8E24AA", "#7B1FA2", "#6A1B9A", "#4A148C",
  # 21-30: Shades of brown for M_SUM species
  "#fff4e3", "#fee0b6", "#fdd697", "#fdb863", "#f29842",
  "#e08214", "#cc6d0d", "#b35806", "#a34906", "#7f3b08"
)

# Step 6: Map values to color
WTT_species_df <- WTT_species_df %>%
  mutate(color = case_when(
    value >= 1  & value <= 30 ~ abandon_palette[value],
    TRUE ~ "#d3d3d3"
  ))

# Step 7: Map
WTT_species_map <- ggplot() +
  geom_tile(data = WTT_species_df, aes(x = x, y = y, fill = color)) +
  scale_fill_identity() +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )  +
  labs(
    title = "Social Benefits through increased ecotourism"
  ) +
  theme(
    plot.title = element_text(
      size = 18,        # font size (increase/decrease as needed)
      face = "bold",    # makes it bold
      hjust = 0.5       # centers the title
    ))
 
# Step 8: Save
ggsave(
  file.path(output_dir, 'WTT_with_species_binary_map_260324.png'),
  plot = WTT_species_map,
  width = 26,
  height = 25,
  units = 'cm',
  dpi = 300
)

# LEGEND SECTION ------------------------------------------------------------

# Step 9: Build the legend data frame
legend_df_WTT <- expand.grid(
  base_level = 1:10,
  expansion = c("No species gained", "P_SUM gained", "M_SUM gained")
)

# Step 10: Define 30-class palette

legend_palette <- c(
  # 1-10: Shades of grey for no abandonment
  "#FAFAFA", "#F0F0F0", "#E0E0E0", "#D6D6D6", "#C4C4C4",
  "#B0B0B0", "#9E9E9E", "#7D7D7D", "#616161", "#424242",
  # 11-20: Shades of purple for P_SUM species
  "#F3E5F5", "#E1BEE7", "#CE93D8", "#BA68C8", "#AB47BC",
  "#9C27B0", "#8E24AA", "#7B1FA2", "#6A1B9A", "#4A148C",
  # 21-30: Shades of brown for M_SUM species
  "#fff4e3", "#fee0b6", "#fdd697", "#fdb863", "#f29842",
  "#e08214", "#cc6d0d", "#b35806", "#a34906", "#7f3b08"
)

# Step 11: Map the correct colors
legend_df_WTT <- legend_df_WTT %>%
  mutate(
    expansion = recode(
      expansion,
      "No species gained" = "Species absent",
      "P_SUM gained" = "Species gained (proba > 0.5)",
      "M_SUM gained" = "No new species"
    ),
    color = case_when(
      expansion == "Species absent" ~ legend_palette[base_level],
      expansion == "Species gained (proba > 0.5)" ~ legend_palette[base_level + 10],
      expansion == "No new species" ~ legend_palette[base_level + 20]
    )
  )

# Step 12: Use rounded NPP quantile labels
WTT_legend_limits <- round(WTT_breaks)

# Step 13: Plot the legend
WTT_species_legend <- ggplot(legend_df_WTT, aes(x = base_level, y = expansion, fill = color)) +
  geom_tile() +
  scale_fill_identity() +
  scale_x_continuous(
    breaks = seq(0.5, 10.5, by = 1),
    labels = WTT_legend_limits,
    position = 'top'
  ) +
  labs(
    x = "Number of people travelling (count/km2)",
    y = "Species expansion status"
  ) +
  theme_minimal() +
  theme(
    axis.title.y = element_text(size = 14, angle = 0, vjust = 1.6, hjust = 1, margin = margin(r = -80)),
    axis.text = element_text(size = 11),
    axis.text.x = element_text(size = 11, angle = 90),
    axis.text.x.top = element_text(size = 11, hjust = 0),
    axis.title.x.top = element_text(size = 14, margin = margin(b = 20)),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(10, 10, 10, 100)
  ) +
  coord_fixed(ratio = 0.5)

# Step 14: Display
plot(WTT_species_legend)

ggsave(
  file.path(output_dir, 'WTT_with_species_binary_legend_260324.png'),
  plot = WTT_species_legend,
  width = 24, height = 10, units = 'cm', dpi = 300
)

# The figure in the article is built upon the maps and legend as separate .png files, here are two supplementary steps to have both together

############ Combine map and legend

# Combine the fixed legend with the map
final_plot <- cowplot::plot_grid(
  WTT_species_map,
  WTT_species_legend,
  ncol = 1, 
  rel_heights = c(5, 1)
)

# Save
ggsave(
  file.path(output_dir, "WTT_with_species_binary_map_with_legend_260324_2.png"),
  plot = final_plot,
  width = 40,
  height = 25,
  units = "cm",
  dpi = 300
)

```

Social conflicts (livestock density x (carnivores current distribution and potential distribution))

```{r}

# Bivariate mapping: binary species expansion x 10 levels of people travelling (wolf & bear only)

# Step 0: Remove overlap between current and proba
M_sub <- M_CN_UA_binary - P_CN_UA_threshold
M_sub <- ifel(M_sub == -1, 0, M_sub)

# Step 1: Livestock and classify into 10 quantiles
livestock_ceil <- ceiling(livestock_sum)
livestock_breaks <- quantile(values(livestock_ceil), probs = seq(1, 0, length.out = 11), na.rm = TRUE)

# Step 2: Create classification matrix with bins 10 → 1, but keep ranges increasing
reversed_class_matrix <- matrix(c(
  livestock_breaks[11] - 0.00001, livestock_breaks[10], 10,
  livestock_breaks[10], livestock_breaks[9], 9,
  livestock_breaks[9],  livestock_breaks[8], 8,
  livestock_breaks[8],  livestock_breaks[7], 7,
  livestock_breaks[7],  livestock_breaks[6], 6,
  livestock_breaks[6],  livestock_breaks[5], 5,
  livestock_breaks[5],  livestock_breaks[4], 4,
  livestock_breaks[4],  livestock_breaks[3], 3,
  livestock_breaks[3],  livestock_breaks[2], 2,
  livestock_breaks[2],  livestock_breaks[1], 1
), ncol = 3, byrow = TRUE)
livestock_classified_10 <- classify(livestock_ceil, rcl = reversed_class_matrix)
plot(livestock_classified_10)

# Step 3: Encode states
livestock_M <- ifel(M_sub == 1 & P_CN_UA_threshold == 0, livestock_classified_10 + 20, NA)
livestock_P <- ifel(P_CN_UA_threshold == 1 & M_sub == 0, livestock_classified_10 + 10, NA)
livestock_none <- ifel(M_sub == 0 & P_CN_UA_threshold == 0, livestock_classified_10, NA)

# Step 4: Merge the rasters
livestock_combined <- merge(livestock_none, livestock_P, livestock_M)
plot(livestock_combined)

# Step 5: Convert to dataframe
livestock_df <- as.data.frame(livestock_combined, xy = TRUE)
colnames(livestock_df) <- c("x", "y", "value")

# Step 6: Define 30-class palette
abandon_palette <- c(
  # 1-10: Shades of grey for no abandonment
  "#FAFAFA", "#F0F0F0", "#E0E0E0", "#D6D6D6", "#C4C4C4",
  "#B0B0B0", "#9E9E9E", "#7D7D7D", "#616161", "#424242",
  # 11-20: Shades of purple for P_SUM species
  "#F3E5F5", "#E1BEE7", "#CE93D8", "#BA68C8", "#AB47BC",
  "#9C27B0", "#8E24AA", "#7B1FA2", "#6A1B9A", "#4A148C",
  # 21-30: Shades of brown for M_SUM species
  "#fff4e3", "#fee0b6", "#fdd697", "#fdb863", "#f29842",
  "#e08214", "#cc6d0d", "#b35806", "#a34906", "#7f3b08"
)

# Step 7: Assign colors
livestock_df <- livestock_df %>%
  mutate(color = case_when(
    value >= 1  & value <= 30 ~ abandon_palette[value],
    TRUE ~ "#d3d3d3"
  ))

# Step 8: Map
livestock_map <- ggplot() +
  geom_tile(data = livestock_df, aes(x = x, y = y, fill = color)) +
  scale_fill_identity() +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2) +
  coord_sf(xlim = c(2635973, 6525973), ylim = c(1448659, 5415659), expand = FALSE) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )  +
  labs(
    title = "Wildlife expansion and low livestock density"
  ) +
  theme(
    plot.title = element_text(
      size = 18,        # font size (increase/decrease as needed)
      face = "bold",    # makes it bold
      hjust = 0.5       # centers the title
    ))

# Step 9: Save

ggsave(
  file.path(output_dir, 'livestock_wolf_bear_species_map_260324.png'),
  plot = livestock_map,
  width = 26, height = 25, units = 'cm', dpi = 300
)

# LEGEND SECTION ------------------------------------------------------------

# Step 10: Build legend df
legend_df <- expand.grid(
  base_level = 1:10,
  expansion = c("No species gained", "P_SUM gained", "M_SUM gained")
)

# Step 11: Recode and assign palette
legend_palette <- abandon_palette
legend_df <- legend_df %>%
  mutate(
    expansion = recode(
      expansion,
      "No species gained" = "Species absent",
      "P_SUM gained" = "Species gained (proba > 0.5)",
      "M_SUM gained" = "No new species"
    ),
    color = case_when(
      expansion == "Species absent" ~ legend_palette[base_level],
      expansion == "Species gained (proba > 0.5)" ~ legend_palette[base_level + 10],
      expansion == "No new species" ~ legend_palette[base_level + 20]
    )
  )

livestock_legend_limits <- round(livestock_breaks)

# Step 12: Plot legend
livestock_legend <- ggplot(legend_df, aes(x = base_level, y = expansion, fill = color)) +
  geom_tile() +
  scale_fill_identity() +
  scale_x_continuous(
    breaks = seq(0.5, 10.5, 1),
    labels = livestock_legend_limits,
    position = 'top'
  ) +
  labs(
    x = "Livestock (count/km2)",
    y = "Species expansion status"
  ) +
  theme_minimal() +
  theme(
    axis.title.y = element_text(size = 14, angle = 0, vjust = 1.6, hjust = 1, margin = margin(r = -80)),
    axis.text = element_text(size = 11),
    axis.text.x = element_text(size = 11, angle = 90),
    axis.text.x.top = element_text(size = 11, hjust = 0),
    axis.title.x.top = element_text(size = 14, margin = margin(b = 20)),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(10, 10, 10, 100)
  ) +
  coord_fixed(ratio = 0.5)

# Step 13: Display
plot(livestock_legend)

# Step 14: Save
ggsave(
  file.path(output_dir, 'livestock_wolf_bear_species_legend_260324.png'),
  plot = livestock_legend,
  width = 24, height = 10, units = 'cm', dpi = 300
)

# The figure in the article is built upon the maps and legend as separate .png files, here are two supplementary steps to have both together

############ Combine map and legend

# Combine the fixed legend with the map
final_plot <- plot_grid(livestock_map, livestock_legend, ncol = 2)

# Save combined
ggsave(
  file.path(output_dir, "livestock_wolf_bear_species_map_with_legend_260324.png"),
  plot = final_plot,
  width = 40, height = 25, units = "cm", dpi = 300
)

```

## IV. Vizualising overlap of the four benefits (top 80% of each)

# 1. Determining when benefits overlap

Get the top 80% of each benefits

```{r}

# Function to compute binary raster for top (80%) bins x binary mask, preserving NA
extract_top80_bins <- function(classified_raster, binary_mask, top_bins = c(3, 4, 5, 6, 7, 8, 9, 10)) {
  result <- ifel(is.na(classified_raster) | is.na(binary_mask), NA,
                 ifel(classified_raster %in% top_bins & binary_mask == 1, 1, 0))
  return(result)
}

# Apply function to each pair
carbon_top80_stack <- extract_top80_bins(pot_carbon_classified, abandonment_masked)
cv_top80_stack  <- extract_top80_bins(cv_classified, abandonment_masked)
wtt_top80_stack <- extract_top80_bins(WTT_classified, P_SUM_threshold)
livestock_top80_stack <- extract_top80_bins(livestock_classified_10, P_CN_UA_threshold)

# Stack for inspection or export
top80_stacks <- c(carbon_top80_stack, cv_top80_stack, wtt_top80_stack, livestock_top80_stack)

par(mfrow = c(2,2))
plot(carbon_top80_stack)
plot(cv_top80_stack)
plot(wtt_top80_stack)
plot(livestock_top80_stack)

# Export to the correct output directory
writeRaster(top80_stacks, filename = file.path(output_dir, "rewilding_opportunity_top80_260324.tif"), overwrite = TRUE)

# NPP Top 80%
png(file.path(output_dir, "carbon_top80_map.png"), width = 800, height = 800, res = 150)
plot(carbon_top80_stack, main = "Top 80% NPP × Abandonment")
dev.off()

# CV Top 80%
png(file.path(output_dir, "cv_top80_map.png"), width = 800, height = 800, res = 150)
plot(cv_top80_stack, main = "Top 80% CV × Abandonment")
dev.off()

# WTT Top 80%
png(file.path(output_dir, "wtt_top80_map.png"), width = 800, height = 800, res = 150)
plot(wtt_top80_stack, main = "Top 80% WTT × Expansion")
dev.off()

# Livestock Top 80%
png(file.path(output_dir, "livestock_top80_map.png"), width = 800, height = 800, res = 150)
plot(livestock_top80_stack, main = "Top 80% Livestock × Expansion")
dev.off()

```

Determining when the four overlap and when they never overlap

```{r}

# Load the four binary top 80% raster layers
carbon80  <- carbon_top80_stack
cv80      <- cv_top80_stack
wtt80     <- wtt_top80_stack
livestock80 <- livestock_top80_stack

### When they never overlapp

# Function to isolate exclusive presence (1) while others are 0
only_one <- function(target, others) {
  result <- ifelse(target == 1 & rowSums(others) == 0, 1, 0)
  result[is.na(target) | apply(is.na(others), 1, any)] <- NA
  return(result)
}

# Convert to matrices for row-wise operations
# Combine other layers excluding each target
others_cv80 <- c(carbon80, wtt80, livestock80)
others_carbon80 <- c(cv80, wtt80, livestock80)
others_wtt80 <- c(carbon80, cv80, livestock80)
others_livestock80 <- c(carbon80, cv80, wtt80)

# Compute exclusive layers
carbon_only80 <- only_one(values(carbon80), values(others_carbon80))
cv_only80 <- only_one(values(cv80), values(others_cv80))
wtt_only80 <- only_one(values(wtt80), values(others_wtt80))
livestock_only80 <- only_one(values(livestock80), values(others_livestock80))

# Assign back to raster objects
carbon_only_r80 <- setValues(carbon80, carbon_only80)
cv_only_r80 <- setValues(cv80, cv_only80)
wtt_only_r80 <- setValues(wtt80, wtt_only80)
livestock_only_r80 <- setValues(livestock80, livestock_only80)

# Combine into a stack
exclusive_stack80 <- c(carbon_only_r80, cv_only_r80, wtt_only_r80, livestock_only_r80)
names(exclusive_stack80) <- c("carbon_only80", "cv_only80", "wtt_only80", "livestock_only80")

# Plot or export
plot(exclusive_stack80)
writeRaster(exclusive_stack80, filename = file.path(output_dir, "exclusive_overlap_stack80_260324.tif"), overwrite = TRUE)

```

No overlap

```{r}

plot(exclusive_stack80)
names(exclusive_stack80)

# Step 1: Assign unique values
carbon_map <- ifel(exclusive_stack80$carbon_only80 == 1, 1, NA)
cv_map <- ifel(exclusive_stack80$cv_only80 == 1, 2, NA)
wtt_map <- ifel(exclusive_stack80$wtt_only80 == 1, 3, NA)
livestock_map <- ifel(exclusive_stack80$livestock_only80 == 1, 4, NA)

# Add background category = 0 where all are NA
all_na <- is.na(carbon_map) & is.na(cv_map) & is.na(wtt_map) &
          is.na(livestock_map)

background_map <- ifel(all_na, 0, NA)  # gray background

# Step 2: Combine layers (priority-based, highest takes precedence)
combined_map <- merge(
  background_map,
  carbon_map, cv_map, wtt_map, livestock_map
)

combined_map_masked <- mask(combined_map, inland_raster)

# Step 3: Convert to dataframe for ggplot
map_df <- as.data.frame(combined_map_masked, xy = TRUE)
colnames(map_df)[3] <- "category"

# Remove NA rows only now
map_df <- map_df[!is.na(map_df$category), ]

# Step 4: Define custom color palette and labels
category_colors <- c(
  "0" = "#ECECEC",  # gray background
  "1" = "#FFD128",
  "2" = "#43C245",
  "3" = "#7DDEEE",
  "4" = "#DB7DEE"
)

category_labels <- c(
  "0" = "No opportunity",
  "1" = "High mitigation",
  "2" = "High adaptation",
  "3" = "High social benefit",
  "4" = "Low conflict"
)

# Saving for the later map
map_df_opp_single_80 <- map_df

# Step 5: Plot with ggplot + borders and coastlines
map_opportunities_one <- ggplot(map_df, aes(x = x, y = y, fill = factor(category))) +
  geom_tile() +
  scale_fill_manual(
    values = category_colors,
    labels = category_labels,
    name = "Single rewilding opportunity"
  ) +
  # Add country borders and coastlines
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),  # adapte les limites si nécessaire
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal()

ggsave(
  file.path(output_dir, 'no_overlap_stack80_260324.png'),
  plot = map_opportunities_one,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

Determining when two benefits overlap

```{r}

### When two layers overlap

# Ensure NA-safe pairwise overlap
binary_overlap <- function(x, y, others) {
  # Condition: x == 1 & y == 1 & all others == 0
  overlap <- (x == 1) & (y == 1)
  exclude <- Reduce("+", lapply(others, function(z) z == 1))
  result <- ifel(overlap & (exclude == 0), 1, 0)
  result[is.na(x) | is.na(y) | Reduce("|", lapply(others, is.na))] <- NA
  return(result)
}

# Compute 2-way overlaps
carbon_cv80        <- binary_overlap(carbon80, cv80, list(wtt80, livestock80))
carbon_wtt80      <- binary_overlap(carbon80, wtt80, list(cv80, livestock80))
carbon_livestock80 <- binary_overlap(carbon80, livestock80, list(cv80, wtt80))
cv_wtt80        <- binary_overlap(cv80, wtt80, list(carbon80, livestock80))
cv_livestock80  <- binary_overlap(cv80, livestock80, list(carbon80, wtt80))
wtt_livestock80 <- binary_overlap(wtt80, livestock80, list(carbon80, cv80))

# Stack results
overlap_stack80 <- c(carbon_cv80, carbon_wtt80, carbon_livestock80, cv_wtt80, cv_livestock80, wtt_livestock80)
names(overlap_stack80) <- c("carbon_cv80", "carbon_wtt80", "carbon_livestock80", "cv_wtt80", "cv_livestock80", "wtt_livestock80")

# Plot or save
plot(overlap_stack80)

# Step 1: Assign unique values
carbon_cv_map80 <- ifel(overlap_stack80$carbon_cv80 == 1, 1, NA)
carbon_wtt_map80 <- ifel(overlap_stack80$carbon_wtt80 == 1, 2, NA)
carbon_livestock_map80 <- ifel(overlap_stack80$carbon_livestock80 == 1, 3, NA)
cv_wtt_map80 <- ifel(overlap_stack80$cv_wtt80 == 1, 4, NA)
cv_livestock_map80 <- ifel(overlap_stack80$cv_livestock80 == 1, 5, NA)
wtt_livestock_map80 <- ifel(overlap_stack80$wtt_livestock80 == 1, 6, NA)

# Add background category = 0 where all are NA
all_na <- is.na(carbon_cv_map80) & is.na(carbon_wtt_map80) & is.na(carbon_livestock_map80) &
          is.na(cv_wtt_map80) & is.na(cv_livestock_map80) & is.na(wtt_livestock_map80)

background_map <- ifel(all_na, 0, NA)  # gray background

# Step 2: Combine layers (priority-based, highest takes precedence)
combined_map <- merge(
  background_map,
  carbon_cv_map80, carbon_wtt_map80, carbon_livestock_map80, cv_wtt_map80, cv_livestock_map80, wtt_livestock_map80  
)

combined_map_masked <- mask(combined_map, inland_raster)

# Step 3: Convert to dataframe for ggplot
map_df <- as.data.frame(combined_map_masked, xy = TRUE)
colnames(map_df)[3] <- "category"

# Remove NA rows only now
map_df <- map_df[!is.na(map_df$category), ]

# Step 4: Define custom color palette and labels
category_colors <- c(
  "0" = "#ECECEC",  # gray background
  "1" = "#D47000",
  "2" = "#00FFE4",
  "3" = "#6C00FF",
  "4" = "#2AFF00",
  "5" = "#0059FF",
  "6" = "#EC01FF"
)

category_labels <- c(
  "0" = "No opportunity",
  "1" = "High mitigation & adaptation",
  "2" = "High mitigation & social benefit",
  "3" = "High mitigation & low conflict",
  "4" = "High adaptation & social benefit",
  "5" = "High adaptation & low conflict",
  "6" = "High social benefit & low conflict"
)

# Saving for the later map
map_df_opp_two_80 <- map_df

# Step 5: Plot with ggplot + borders and coastlines
map_opportunities_two <- ggplot(map_df, aes(x = x, y = y, fill = factor(category))) +
  geom_tile() +
  scale_fill_manual(
    values = category_colors,
    labels = category_labels,
    name = "Two rewilding opportunities"
  ) +
  # Add country borders and coastlines
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),  # adapte les limites si nécessaire
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal()

ggsave(
  file.path(output_dir, 'two_overlap_stack80_260324.png'),
  plot = map_opportunities_two,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

Determining when three benefits overlap

```{r}

### When three layers overlap

# Function to compute 3-way overlap while the 4th is 0
three_way_overlap <- function(a, b, c, not_d) {
  combo <- (a == 1) & (b == 1) & (c == 1) & (not_d == 0)
  combo[is.na(a) | is.na(b) | is.na(c) | is.na(not_d)] <- NA
  return(combo)
}

# Compute all 4 possible 3-way overlaps (excluding one each time)
overlap_carbon_cv_wtt80     <- three_way_overlap(carbon80, cv80, wtt80, livestock80)
overlap_carbon_cv_livest80 <- three_way_overlap(carbon80, cv80, livestock80, wtt80)
overlap_carbon_wtt_livest80 <- three_way_overlap(carbon80, wtt80, livestock80, cv80)
overlap_cv_wtt_livest80  <- three_way_overlap(cv80, wtt80, livestock80, carbon80)

# Combine into a stack
overlap_3_stack80 <- c(
  overlap_carbon_cv_wtt80,
  overlap_carbon_cv_livest80,
  overlap_carbon_wtt_livest80,
  overlap_cv_wtt_livest80
)
names(overlap_3_stack80) <- c("carbon_cv_wtt80", "carbon_cv_livest80", "carbon_wtt_livest80", "cv_wtt_livest80")

plot(overlap_3_stack80)
names(overlap_3_stack80)

# Step 1: Assign unique values
carbon_cv_wtt_map80 <- ifel(overlap_3_stack80$carbon_cv_wtt80 == 1, 1, NA)
carbon_cv_livest_map80 <- ifel(overlap_3_stack80$carbon_cv_livest80 == 1, 2, NA)
carbon_wtt_livest_map80 <- ifel(overlap_3_stack80$carbon_wtt_livest80 == 1, 3, NA)
cv_wtt_livest_map80 <- ifel(overlap_3_stack80$cv_wtt_livest80 == 1, 4, NA)

# Add background category = 0 where all are NA
all_na <- is.na(carbon_cv_wtt_map80) & is.na(carbon_cv_livest_map80) & is.na(carbon_wtt_livest_map80) & is.na(cv_wtt_livest_map80)

background_map <- ifel(all_na, 0, NA)  # gray background

# Step 2: Combine layers (priority-based, highest takes precedence)
combined_map <- merge(
  background_map,
  carbon_cv_wtt_map80, carbon_cv_livest_map80, carbon_wtt_livest_map80, cv_wtt_livest_map80
)

combined_map_masked <- mask(combined_map, inland_raster)

# Step 3: Convert to dataframe for ggplot
map_df <- as.data.frame(combined_map_masked, xy = TRUE)
colnames(map_df)[3] <- "category"

# Remove NA rows only now
map_df <- map_df[!is.na(map_df$category), ]

# Step 4: Define custom color palette and labels
category_colors <- c(
  "0" = "#ECECEC",  # gray background
  "1" = "#000FFF",
  "2" = "#00C112", 
  "3" = "#BA01FF",
  "4" = "#01FFEC"
)

category_labels <- c(
  "0" = "No opportunity",
  "1" = "High mitigation, high adaptation & high social benefit",
  "2" = "High mitigation, high adaptation & low conflict",
  "3" = "High mitigation, high social benefit & low conflict",
  "4" = "High adaptation, high social benefit & low conflict"
)

# Saving for the later map
map_df_opp_three_80 <- map_df 

# Step 5: Plot with ggplot + borders and coastlines
map_opportunities_three <- ggplot(map_df, aes(x = x, y = y, fill = factor(category))) +
  geom_tile() +
  scale_fill_manual(
    values = category_colors,
    labels = category_labels,
    name = "Three rewilding opportunities"
  ) +
  # Add country borders and coastlines
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),  # adapte les limites si nécessaire
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal()

ggsave(
  file.path(output_dir, 'three_overlap_stack80_260324.png'),
  plot = map_opportunities_three,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

Determining when the four benefits overlap

```{r}

### When four layers overlap
combined_top80_all1 <- ifel(
  is.na(carbon80) | is.na(cv80) |
  is.na(wtt80) | is.na(livestock80),
  NA,
  ifel(carbon80 == 1 & cv80 == 1 &
       wtt80 == 1 & livestock80 == 1,
       1, 0)
)

plot(combined_top80_all1)
names(combined_top80_all1)

combined_top80_all1_masked <- mask(combined_top80_all1, inland_raster)

# Step 1: Convert to dataframe
map_df <- as.data.frame(combined_top80_all1_masked, xy = TRUE)
colnames(map_df)[3] <- "category"

# Remove NA rows only
map_df <- map_df[!is.na(map_df$category), ]

map_df <- map_df %>%
  filter(x >= 2635973, x <= 6525973, y >= 1448659, y <= 5415659)

# Step 2: Define color palette and labels
category_colors <- c(
  "0" = "#ECECEC",  # gray background
  "1" = "#01FF05"   # you can pick another color
)

category_labels <- c(
  "0" = "No opportunity",
  "1" = "High mitigation, high adaptation &\n high social benefit & low conflict"
)

# Saving for the later map
map_df_opp_four_80 <- map_df

# Step 3: Plot
map_opportunities_four <- ggplot(map_df, aes(x = x, y = y, fill = factor(category))) +
  geom_tile() +
  scale_fill_manual(
    values = category_colors,
    labels = category_labels,
    name = "Four rewilding opportunities"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal()

ggsave(
  file.path(output_dir, 'four_overlap_stack80_260324.png'),
  plot = map_opportunities_four,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

Map the number of overlaps

```{r}

# Step 1: Keep 0 when 0, retain all values
map_df_opp_single_80$category <- ifelse(map_df_opp_single_80$category != 0, 1, 0)
map_df_opp_two_80$category    <- ifelse(map_df_opp_two_80$category != 0, 2, 0)
map_df_opp_three_80$category  <- ifelse(map_df_opp_three_80$category != 0, 3, 0)
map_df_opp_four_80$category   <- ifelse(map_df_opp_four_80$category != 0, 4, 0)

# Step 2: Keep only rows where category != 0 (to avoid double-counting), retain ocean as NA
df1 <- map_df_opp_single_80[map_df_opp_single_80$category != 0, ]
df2 <- map_df_opp_two_80[map_df_opp_two_80$category != 0, ]
df3 <- map_df_opp_three_80[map_df_opp_three_80$category != 0, ]
df4 <- map_df_opp_four_80[map_df_opp_four_80$category != 0, ]

# Step 3: Merge priority-wise (4 → 3 → 2 → 1)
map_df <- rbind(df1, df2, df3, df4)
map_df <- map_df[order(map_df$x, map_df$y, -map_df$category), ]  # prioritize higher category
map_df <- map_df[!duplicated(map_df[, c("x", "y")]), ]            # remove duplicates, keep highest

# Step 4: Rebuild NA + 0 base from any of the original full maps
base_template <- map_df_opp_single_80[, c("x", "y")]
map_df_final <- merge(base_template, map_df, by = c("x", "y"), all.x = TRUE)

# Step 5: Replace NA in category with 0 (only for land areas that had no opportunities)
# Note: leave real NA (ocean) untouched
map_df_final$category <- ifelse(
  is.na(map_df_final$category) & !is.na(map_df_opp_single_80$category),
  0,
  map_df_final$category
)

# Step 6: Factor conversion for plotting
map_df_final$category <- factor(map_df_final$category, levels = 0:4)

# Step 4: Define custom color palette and labels
category_colors <- c(
  "0" = "#FFFFFF",  # white background
  "1" = "#A9BF00",  
  "2" = "#FCCA3F",  
  "3" = "#FC3F3F",  
  "4" = "#7363FF"   
)

category_labels <- c(
  "0" = "No opportunity",
  "1" = "Single opportunity",
  "2" = "Two opportunities",
  "3" = "Three opportunities",
  "4" = "Four opportunities"
)

# Step 5: Plot base map
map_opportunities_number <- ggplot(map_df_final, aes(x = x, y = y, fill = factor(category))) +
  geom_tile() +
  scale_fill_manual(
    values = category_colors,
    labels = category_labels,
    name = "Rewilding potential"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal()

ggsave(
  file.path(output_dir, 'number_of_overlap_stack80_260324.png'),
  plot = map_opportunities_number,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

# 2. Determining benefits extent: distinguishing overlapping extent and exclusive extent 

Climate change mitigation

```{r}

######## Carbon
# Whole extent
plot(carbon80)
# Extent alone
plot(exclusive_stack80$carbon_only80)

# Step 1: Reclassify rasters into 0/1
r_whole <- carbon80
r_excl  <- exclusive_stack80$carbon_only80

# Step 2: Combine with priority to exclusive
# Start with whole extent
r_combined <- ifel(r_whole == 1, 1, 0)  # 1 = whole
# Overwrite with exclusive where present
r_combined <- ifel(r_excl == 1, 2, r_combined)  # 2 = exclusive

# Step 3: Convert to data frame for ggplot
df_combined <- as.data.frame(r_combined, xy = TRUE)
names(df_combined)[3] <- "category"
df_combined$category <- factor(df_combined$category, levels = c(0, 1, 2))

# Step 4: Define colors and labels
combined_colors <- c(
  "0" = "#FFFFFF",   # white
  "1" = "#0090FF",   # light blue = whole extent
  "2" = "#D10000"    # dark blue = exclusive
)

combined_labels <- c(
  "1" = "Overlapping extent",
  "2" = "Exclusive extent"
)

# Step 5: Plot
map_combined_carbon <- ggplot(df_combined, aes(x = x, y = y, fill = category)) +
  geom_tile() +
  scale_fill_manual(
    values = combined_colors,
    breaks = c("1", "2"),    # show only meaningful classes
    labels = combined_labels[c("1", "2")],
    name = "CC mitigation opportunity"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "Climate change mitigation opportunity (top 80%)")

ggsave(
  file.path(output_dir, 'opportunities80_mitigation_260327.png'),
  plot = map_combined_carbon,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

Climate change adaptation

```{r}

# Whole extent
plot(cv_top80_stack)
# Extent alone
plot(exclusive_stack80$cv_only80)

# Step 1: Reclassify rasters into 0/1
r_whole <- cv_top80_stack
r_excl  <- exclusive_stack80$cv_only80

# Step 2: Combine with priority to exclusive
# Start with whole extent
r_combined <- ifel(r_whole == 1, 1, 0)  # 1 = whole
# Overwrite with exclusive where present
r_combined <- ifel(r_excl == 1, 2, r_combined)  # 2 = exclusive

# Step 3: Convert to data frame for ggplot
df_combined <- as.data.frame(r_combined, xy = TRUE)
names(df_combined)[3] <- "category"
df_combined$category <- factor(df_combined$category, levels = c(0, 1, 2))

# Step 4: Define colors and labels
combined_colors <- c(
  "0" = "#FFFFFF",   # white
  "1" = "#0090FF",   # light blue = whole extent
  "2" = "#D10000"    # dark blue = exclusive
)

combined_labels <- c(
  "1" = "Overlapping extent",
  "2" = "Exclusive extent"
)

# Step 5: Plot
map_combined_cv <- ggplot(df_combined, aes(x = x, y = y, fill = category)) +
  geom_tile() +
  scale_fill_manual(
    values = combined_colors,
    breaks = c("1", "2"),    # show only meaningful classes
    labels = combined_labels[c("1", "2")],
    name = "CC adaptation opportunity"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "Climate change adaptation opportunity (top 80%)")

ggsave(
  file.path(output_dir, 'opportunities80_adaptation_260327.png'),
  plot = map_combined_cv,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

Social benefits

```{r}

# Whole extent
plot(wtt_top80_stack)
# Extent alone
plot(exclusive_stack80$wtt_only80)

# Step 1: Reclassify rasters into 0/1
r_whole <- wtt_top80_stack
r_excl  <- exclusive_stack80$wtt_only80

# Step 2: Combine with priority to exclusive
# Start with whole extent
r_combined <- ifel(r_whole == 1, 1, 0)  # 1 = whole
# Overwrite with exclusive where present
r_combined <- ifel(r_excl == 1, 2, r_combined)  # 2 = exclusive

# Step 3: Convert to data frame for ggplot
df_combined <- as.data.frame(r_combined, xy = TRUE)
names(df_combined)[3] <- "category"
df_combined$category <- factor(df_combined$category, levels = c(0, 1, 2))

# Step 4: Define colors and labels
combined_colors <- c(
  "0" = "#FFFFFF",   # white
  "1" = "#0090FF",   # light blue = whole extent
  "2" = "#D10000"    # dark blue = exclusive
)

combined_labels <- c(
  "1" = "Overlapping extent",
  "2" = "Exclusive extent"
)

# Step 5: Plot
map_combined_wtt <- ggplot(df_combined, aes(x = x, y = y, fill = category)) +
  geom_tile() +
  scale_fill_manual(
    values = combined_colors,
    breaks = c("1", "2"),    # show only meaningful classes
    labels = combined_labels[c("1", "2")],
    name = "Ecotourism opportunity"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "Ecotourism opportunity (top 80%)")

ggsave(
  file.path(output_dir, 'opportunities80_wtt_260327.png'),
  plot = map_combined_wtt,
  width = 26, height = 25, units = 'cm', dpi = 300
)


```

Sociale conflicts

```{r}

# Whole extent
plot(livestock_top80_stack)
# Extent alone
plot(exclusive_stack80$livestock_only80)

# Step 1: Reclassify rasters into 0/1
r_whole <- livestock_top80_stack
r_excl  <- exclusive_stack80$livestock_only80

# Step 2: Combine with priority to exclusive
# Start with whole extent
r_combined <- ifel(r_whole == 1, 1, 0)  # 1 = whole
# Overwrite with exclusive where present
r_combined <- ifel(r_excl == 1, 2, r_combined)  # 2 = exclusive

# Step 3: Convert to data frame for ggplot
df_combined <- as.data.frame(r_combined, xy = TRUE)
names(df_combined)[3] <- "category"
df_combined$category <- factor(df_combined$category, levels = c(0, 1, 2))

# Step 4: Define colors and labels
combined_colors <- c(
  "0" = "#FFFFFF",   # white
  "1" = "#0090FF",   # light blue = whole extent
  "2" = "#D10000"    # dark blue = exclusive
)

combined_labels <- c(
  "1" = "Overlapping extent",
  "2" = "Exclusive extent"
)

# Step 5: Plot
map_combined_livestock <- ggplot(df_combined, aes(x = x, y = y, fill = category)) +
  geom_tile() +
  scale_fill_manual(
    values = combined_colors,
    breaks = c("1", "2"),    # show only meaningful classes
    labels = combined_labels[c("1", "2")],
    name = "Low conflict opportunity"
  ) +
  geom_sf(data = country_borders, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  geom_sf(data = coastlines, fill = NA, color = "black", size = 0.2, inherit.aes = FALSE) +
  coord_sf(
    xlim = c(2635973, 6525973),
    ylim = c(1448659, 5415659),
    expand = FALSE
  ) +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")
  ) +
  labs(title = "Low livestock-wildlife conflict opportunity (top 80%)")

ggsave(
  file.path(output_dir, 'opportunities80_livestock_260327.png'),
  plot = map_combined_livestock,
  width = 26, height = 25, units = 'cm', dpi = 300
)

```

