---
title: "maps_design"
author: "Magali Weissgerber"
date: "2026-02-04"
output: html_document
---

## Loading libraries

```{r}

library(terra)

library(ggplot2)

library(tidyterra)

library(sf)

library(gridExtra)

library(ggspatial)

library(cowplot)

library(ggtext)

library(xfun)

library(patchwork)

library(dplyr)

library(grid)

```

## Prepare a reference raster and a mask with Europe's borders

```{r}

# Import the raster of human population and of Europe's border to have the reference raster regarding crs, ext and res (ie the population raster is the reference one)

# Import population raster nut ditinguisg inland (should be 0) and oceans (should be NA)

# Step 1: Load the population raster and coastline shapefile
raster_1km_human_population_path <- "S:/Magali_Weissgerber/human_population_scratch/JRC_1K_POP_2018.tif" # Read the 1km resolution raster
raster_1km_human_population <- rast(raster_1km_human_population_path)
plot(raster_1km_human_population)

# Import Europe's borders
inland_shapefile <- st_read("S:/Magali_Weissgerber/Europe frontiers_scratch/NUT3_from_rewilding_folder/Europe_limit_from_nut3_dissolve.shp")
plot(inland_shapefile)

# Check the type of the raster object
class(raster_1km_human_population)

# Check if the raster object is valid
print(raster_1km_human_population)

# Ensure the CRS is consistent
crs(raster_1km_human_population) <- "EPSG:3035"
# Check and reproject the shapefile to match the CRS of the raster
if (crs(inland_shapefile) != crs(raster_1km_human_population)) {
  inland_shapefile <- st_transform(inland_shapefile, crs(raster_1km_human_population))
}

# Step 2: Create a mask from the coastline shapefile
# Rasterize the coastline shapefile to create a mask
inland_raster <- rasterize(vect(inland_shapefile), raster_1km_human_population, field = 1, background = NA)
plot(inland_raster)

# Step 3: Preprocess the population raster
# Replace inland NA values with 0
population_preprocessed <- raster_1km_human_population
values(population_preprocessed)[is.na(values(population_preprocessed)) & !is.na(values(inland_raster))] <- 0
plot(population_preprocessed)

#Rename
raster_1km_human_population <- population_preprocessed

#Export
#output_population_preprocessed <- ("S:/Magali_Weissgerber/Species raster for social benefit of rewilding/population_masked_ocean_240723.tif")
#writeRaster(raster_1km_human_population, output_population_preprocessed, overwrite=TRUE)

```

## Step 1 - determining the expansion probability of large mammals

We are using the wildlife maps from the Rewild Europe Report (Fernández, N., Torres, A., Wolf, F., Quintero, L., & Pereira, H. M. (2020). Boosting Ecological Restoration for a Wilder Europe. Halle-Wittenberg: German Centre for Integrative Biodiversity Research (iDiv) and Martin-Luther-Universität.)

They include Alpine ibex, Brown bear, Chamois, Elk, European bison, European lynx, European wolf, Iberian ibex, Iberian lynx, Pyrenean chamois, Red deer, Reindeer, Roe deer, Wild boar, Wild goat.
We will focus on Brown bear, Elk, European bison, European wolf, European lynx, Iberian lynx, Iberian ibex, Alpine ibex.

We want to know the probability that cells not containing the focus species will gain that species. Here are the different steps.
For each species
1) Occurrences are coded as 0 = species absent, 1 = species present and 9 = species maybe present. We are changing all the 9 into 0.
2) Get the distance of every 0 to the closest 1
3) Calculate probability of colonization (form K/e^-xdistance)
Then
4) Sum the probability (Optional)

A parallel processing was added to rasterize the data directly and change the resolution from 10km to 1km, and calculate distance then.

### Import and check data; change to binary

```{r}

SPECIES_OCC_initial_fishnet <-  st_read("S:/Magali_Weissgerber/Species_dist_occ_scratch/SP_OCCU_190207.shp")

head(SPECIES_OCC_initial_fishnet)
names(SPECIES_OCC_initial_fishnet)

ggplot(SPECIES_OCC_initial_fishnet) + 
  geom_sf(aes(fill = M_CANLUP), color = NA) +
  scale_fill_viridis_c(option = "magma",begin = 0.2)

SPECIES_OCC_binary_fishnet <- SPECIES_OCC_initial_fishnet #new dataset to be modified
  
# Change 9 into 0
  for(j in 4:20)
    set(SPECIES_OCC_binary_fishnet, i = which(SPECIES_OCC_binary_fishnet[[j]] == 9), j = j, value = 0)

#Plot one species to check
ggplot(SPECIES_OCC_binary_fishnet) + 
  geom_sf(aes(fill = M_CANLUP), color = NA) +
  scale_fill_viridis_c(option = "magma",begin = 0.2)

```

### Rasterize, create distance raster and calculate probability of expansion

Calculate probability of colonization (form K/e^-xdistance) (when distance = 0 it means the species is already there so probability is also 0). 
Initially we used 1/exp((log(2)/50) distance) so to have 50% of probability of comeback at 50km. But based on Canis lupus dispersal values we apply rather 1/exp((log(2)/60) distance) (as the cells of the mammals ditribution are 10x10km the probability max is 0.89)

```{r}

# Rasterize the shapefile. It becomes a stack raster with the 8 species of interest

# Define the extent and resolution for the raster
extent_raster <- terra::ext(SPECIES_OCC_binary_fishnet)
res <- 10000  # Define your desired resolution
# Create an empty raster with the defined extent and resolution
r <- terra::rast(extent_raster, resolution = res, crs = crs(SPECIES_OCC_binary_fishnet))
# List of columns to rasterize
columns_to_rasterize <- c("M_CANLUP", "M_ALCALC", "M_BISBON", "M_LYNLYN", "M_URSARC", "M_LYNPAR", "M_CAPIBE", "M_CAPPYR")
# Initialize an empty list to store rasters
rasters <- list()
# Loop through each column and rasterize
for (col in columns_to_rasterize) {
    r <- rasterize(SPECIES_OCC_binary_fishnet, r, field = col)
    rasters[[col]] <- r
}
# Stack all rasters
raster_10km_wildlife_occupancy_stack <- rast(rasters)
plot(raster_10km_wildlife_occupancy_stack)

# Change the resolution of the wildlife raster 

# Check the original raster resolutions
print(res(raster_10km_wildlife_occupancy_stack))
print(res(raster_1km_human_population))

# Resample the 10km raster to 1km resolution using nearest neighbor method
raster_1km_wildlife_occupancy_stack <- resample(raster_10km_wildlife_occupancy_stack, raster_1km_human_population, method = "ngb")

# Check the resampled raster resolution
print(res(raster_1km_wildlife_occupancy_stack))

# Plot raster
plot(raster_1km_wildlife_occupancy_stack)

# Sum the occupancy (for illustrative maps)
raster_1km_wildlife_occupancy_stack$M_sum <- app(raster_1km_wildlife_occupancy_stack, sum)
plot(raster_1km_wildlife_occupancy_stack)

#Export
#output_file_occupancy <- ("S:/Magali_Weissgerber/Species raster for social benefit of rewilding/SP_occupancy_250106.tif")
#writeRaster(raster_1km_wildlife_occupancy_stack, output_file_occupancy, overwrite=TRUE)

# let's produce the distance raster for each species (it can take 7h per species to process)

# 1. Create list of species rasters

species_list <- list(
  CANLUP = raster_1km_wildlife_occupancy_stack$M_CANLUP,
  ALCALC = raster_1km_wildlife_occupancy_stack$M_ALCALC,
  BISBON = raster_1km_wildlife_occupancy_stack$M_BISBON,
  LYNLYN = raster_1km_wildlife_occupancy_stack$M_LYNLYN,
  URSARC = raster_1km_wildlife_occupancy_stack$M_URSARC,
  LYNPAR = raster_1km_wildlife_occupancy_stack$M_LYNPAR,
  CAPIBE = raster_1km_wildlife_occupancy_stack$M_CAPIBE,
  CAPPYR = raster_1km_wildlife_occupancy_stack$M_CAPPYR
)

# 2. Function to process species (no export)

process_species <- function(r, species_name) {
  message("Processing species: ", species_name)

  # Step 1: Clean raster (replace NaN with NA)
  r_clean <- classify(r, matrix(c(NaN, NA), ncol = 2, byrow = TRUE))
  r_clean <- mask(r_clean, r_clean)

  # Step 2: Check unique values
  message("Unique values in cleaned raster:")
  print(unique(values(r_clean)))

  # Step 3: Extract presence cells (value = 1)
  pts <- as.points(r_clean, values = TRUE, na.rm = TRUE)
  pts <- pts[pts[[1]] == 1, ]

  # Step 4: Compute distance raster (km)
  dist_r <- distance(r_clean, pts, unit = "km")

  # Optional plot
  plot(dist_r, main = paste0("Distance Raster: ", species_name))

  # Return list of outputs
  return(list(
    cleaned_raster = r_clean,
    points = pts,
    distance_raster = dist_r
  ))
}

# 3. Function to export a processed species output

export_species <- function(distance_raster, species_name, output_dir) {

  # Build filename with date
  output_file <- file.path(
    output_dir,
    paste0("dist_", species_name, "_", format(Sys.Date(), "%y%m%d"), ".tif")
  )

  writeRaster(distance_raster, output_file, overwrite = TRUE)
  message("Exported: ", output_file)
}

# 4. Apply processing to all species (no export happens here)

processed_results <- mapply(
  FUN  = process_species,
  r    = species_list,
  species_name = names(species_list),
  SIMPLIFY = FALSE
)

# 5. Export only when you choose

output_directory <- "S:/Magali_Weissgerber/Species raster for social benefit of rewilding"

# Example: Export LYNPAR and CAPIBE
export_species(processed_results$LYNPAR$distance_raster, "LYNPAR", output_directory)
export_species(processed_results$CAPIBE$distance_raster, "CAPIBE", output_directory)

# Next import 

# 1. List species names

species_names <- c("M_CANLUP", "M_ALCALC", "M_BISBON", "M_LYNLYN", "M_URSARC", "M_LYNPAR", "M_CAPIBE", "M_CAPPYR")
# Directory where files were exported
input_dir <- "S:/Magali_Weissgerber/Species raster for social benefit of rewilding"

# 2. Create a function: import + plot + mask + return

import_and_mask_species <- function(species, input_dir, inland_raster) {
  
  # Find the correct file automatically (matching species name)
  file_pattern <- paste0("dist_", species, "_")
  file_path <- list.files(input_dir, pattern = file_pattern, full.names = TRUE)
  
  if (length(file_path) == 0) {
    stop("No file found for species: ", species)
  }
  
  message("Importing raster for: ", species)
  
  # Import
  r <- rast(file_path)
  plot(r, main = paste("Imported Distance Raster:", species))
  
  # Mask
  r_masked <- mask(r, inland_raster)
  plot(r_masked, main = paste("Masked Distance Raster:", species))
  
  return(r_masked)
}

# 3. Apply this function to all species automatically

masked_rasters <- lapply(
  species_names,
  import_and_mask_species,
  input_dir = input_dir,
  inland_raster = inland_raster
)

# The list will be named for convenience:
names(masked_rasters) <- species_names

# Check results

masked_rasters$LYNPAR
masked_rasters$CAPIBE
masked_rasters$CAPPYR

# Then, calculate probability for expansion (based on distance to 1)

# Step 1: Define the transformation function
probability_function <- function(distance) {
  ifelse(distance == 0, 
         0, 
         1 / exp((log(2) / 60) * distance))
}


# Apply the function to each raster in the stack
species_probability_stack <- app(species_stack, probability_function)

# Rename the rasters: replace "M_" with "P_" in layer names (the distribution fields are called M_CANLUP, etc)
old_names <- names(masked_rasters)               # Get original names
new_names <- gsub("^M_", "P_", old_names)       # Replace "M_" with "P_"
names(species_probability_stack) <- new_names   # Assign new names to result_stack

# Calculate the sum of the 8 species rasters
sum_raster <- sum(species_probability_stack)

# Rename the sum raster
names(sum_raster) <- "P_SUM"

# Add the sum raster to the stack
species_probability_stack_2 <- c(species_probability_stack, sum_raster)

# Plot results
plot(species_probability_stack_2)

#Export
output_file_Proba <- ("S:/Magali_Weissgerber/Species raster for social benefit of rewilding/SP_expansion_probability_250106.tif")
writeRaster(species_probability_stack_2, output_file_Proba, overwrite=TRUE)


```

## Step 2 - determining social benefit

We look at people's willingness-To-Travel (WTT) to see wildlife.
WTT is distributed among people, we use this distribution to determine the relation between distance to a cell and the share of people in the surrounding cells that will travel. Use the complementary lognormal CDF function (1 - CDF) and the density of people in a convolution to determine the amount of people going to one cell.

# Prepare rasters

```{r}

# The human population raster has been pre processes already 
plot(raster_1km_human_population)

# Downscale the current occupancy + probability (10km) raster to match human population (1km)

# Current occupancy + probability

# Check the original raster resolutions
print(res(raster_10km_wildlife_current_probability))
print(res(raster_1km_human_population))

# Resample the 10km raster to 1km resolution to match the 1km raster
raster_1km_wildlife_current_probability_stack<- resample(raster_10km_wildlife_current_probability, raster_1km_human_population, method = "bilinear")

# Check the resampled raster resolution
print(res(raster_1km_wildlife_current_probability_stack))

#Plot raster
plot(raster_1km_wildlife_current_probability_stack)

# Sum the probability (for illustrative maps)
raster_1km_wildlife_current_probability_stack$MP_sum <- app(raster_1km_wildlife_current_probability_stack, sum)
plot(raster_1km_wildlife_current_probability_stack)

#Export
output_file_MP <- ("S:/Magali_Weissgerber/Species raster for social benefit of rewilding/SP_OCCU_current_probability_241212.tif")
writeRaster(raster_1km_wildlife_current_probability_stack, output_file_MP, overwrite=TRUE)

# We look at people's willingness-To-Travel to see wildlife Possibility: we exclude rural areas because they mostly have negative opinions of wildlife comeback. 
# Create a new raster (threshold can be applied now, the current version has no threshold)
raster_1km_human_population_threshold <- raster_1km_human_population

```

# Calculate the number of people travelling based on the WTT

For large carnivores

```{r}

# Step 1: create the kernel matrix

# Step 1.1: Define matrix dimensions and create the distance matrix
nrows <- nrow(raster_1km_human_population_threshold)
ncols <- ncol(raster_1km_human_population_threshold)
x_coords <- matrix(rep(seq(1, ncols), each = nrows), nrow = nrows, ncol = ncols)
y_coords <- matrix(rep(seq(1, nrows), times = ncols), nrow = nrows, ncol = ncols)
center_x <- ncols / 2
center_y <- nrows / 2
distance_matrix_LC <- sqrt((x_coords - center_x)^2 + (y_coords - center_y)^2)

# Step 1.2: Define parameters for the lognormal kernel
mu <- -2.23846
sigma <- 1.000746

# Step 1.3: Create the lognormal kernel matrix
kernel_matrix_LC <- dlnorm(distance_matrix_LC, meanlog = mu, sdlog = sigma)

# Step 1.4: Check the dimensions and a small portion of the matrix
dim(kernel_matrix_LC)
print("Top-left corner of the matrix:")
print(kernel_matrix_LC[1:10, 1:10])  # Displaying the top-left 10x10 cells

#Step 2: create the population matrix (the ocean's 0 were masked and change to NAs previously because convolution was done in the spatial domain, with this faster method using fourier transform the NAs have to be set back to 0 for the matrix, yet I kept it that way)

#Step 2.1: create an identifcal raster and print basic information about the raster
raster_1km_human_population_threshold_2 <- raster_1km_human_population_threshold
print(raster_1km_human_population_threshold_2)
# Summary of the raster to check the range of values
summary(raster_1km_human_population_threshold_2)

#Step 2.2: Check the number of non-zero and NA values
num_zeros <- sum(values(raster_1km_human_population_threshold_2) == 0, na.rm = TRUE)
num_nas <- sum(is.na(values(raster_1km_human_population_threshold_2)))
num_non_zeros <- sum(values(raster_1km_human_population_threshold_2) > 0, na.rm = TRUE)
cat("Number of zero values:", num_zeros, "\n")
cat("Number of NA values:", num_nas, "\n")
cat("Number of non-zero values:", num_non_zeros, "\n")

#Step 2.3: Replace NA values with zeros (if appropriate for your analysis)
raster_1km_human_population_threshold_2[is.na(raster_1km_human_population_threshold_2)] <- 0
plot(raster_1km_human_population_threshold_2)
# Check if NA values are replaced
na_count_after <- sum(is.na(values(raster_1km_human_population_threshold_2)))
print(paste("Number of NA values after replacement:", na_count_after))

#Step 2.4: Get the number of rows and columns from the original raster
nrows <- nrow(raster_1km_human_population_threshold_2)
ncols <- ncol(raster_1km_human_population_threshold_2)

#Step 2.5: Convert the raster to a matrix and reshape it
population_matrix_LC <- matrix(values(raster_1km_human_population_threshold_2), 
                            nrow = nrows, ncol = ncols, byrow = TRUE)
# Check the dimensions and a small portion of the matrix
dim(population_matrix_LC)
print("Top-left corner of the matrix:")
print(population_matrix_LC[1:10, 1:10])  # Displaying the top-left 10x10 cells

# Step 3: Use the 'Fast-Fourier' transform to compute the convolution
convolved_matrix_LC <- convolve_image(population_matrix_LC, kernel_matrix_LC)

# Step 4: convert the matrix to raster

# Step 4.1: Convert the convolved matrix back to raster format using terra's rast function
convolved_raster_LC <- rast(nrows = nrow(raster_1km_human_population_threshold_2),
                                  ncols = ncol(raster_1km_human_population_threshold_2),
                                  xmin = xmin(raster_1km_human_population_threshold_2),
                                  xmax = xmax(raster_1km_human_population_threshold_2),
                                  ymin = ymin(raster_1km_human_population_threshold_2),
                                  ymax = ymax(raster_1km_human_population_threshold_2))

# Step 4.2: Assign the values of the convolved matrix back to the raster
values(convolved_raster_LC) <- convolved_matrix_LC

# Step 4.3: fix crs, extent and resolution
res(raster_1km_human_population_threshold_2)
res(convolved_raster_LC)
crs(raster_1km_human_population_threshold_2)
crs(convolved_raster_LC)
ext(raster_1km_human_population_threshold_2)
ext(convolved_raster_LC)

# Step 4.3: fix crs
crs(convolved_raster_LC) <- crs(raster_1km_human_population_threshold_2)

# Step 5: Plot results
plot(raster_1km_human_population_threshold_2, main = "Original Population Raster")
plot(kernel_matrix_LC, main = "Kernel Raster")
plot(convolved_raster_LC, main = "Convolved Raster with Identity Kernel")

# Step 6: mask the ocean

# Step 6.1: import the inland shapefile
inland_shapefile <- st_read("S:/Magali_Weissgerber/Europe frontiers_scratch/NUT3_from_rewilding_folder/Europe_limit_from_nut3_dissolve.shp")

# Step 6.2: mask the output raster
convolved_raster_masked_LC <- mask(convolved_raster_LC, inland_raster)
plot(convolved_raster_masked_LC)

# Compute quantile breaks (exclude NA values)
q_breaks <- quantile(values(convolved_raster_masked_LC), 
                     probs = seq(0, 1, length.out = 11), 
                     na.rm = TRUE)

# Plot with 10 quantile levels
plot(convolved_raster_masked_LC, 
     breaks = q_breaks, 
     col = hcl.colors(10, "YlOrRd", rev = TRUE), 
     main = "Convolved Raster (Quantile Breaks) large carnivores")

# Step 7: export
#output_file_WTT_masked <- ("S:/Magali_Weissgerber/Species raster for social benefit of rewilding/WTT_LC_raster_masked_240925.tif")
#writeRaster(convolved_raster_masked_LC, output_file_WTT_masked, overwrite=TRUE)



```

For large herbivores

```{r}

# Step 1: create the kernel matrix

# Step 1.1: Define matrix dimensions and create the distance matrix
nrows <- nrow(raster_1km_human_population_threshold)
ncols <- ncol(raster_1km_human_population_threshold)
x_coords <- matrix(rep(seq(1, ncols), each = nrows), nrow = nrows, ncol = ncols)
y_coords <- matrix(rep(seq(1, nrows), times = ncols), nrow = nrows, ncol = ncols)
center_x <- ncols / 2
center_y <- nrows / 2
distance_matrix_LH <- sqrt((x_coords - center_x)^2 + (y_coords - center_y)^2)

# Step 1.2: Define parameters for the lognormal kernel
mu <- -1.50609
sigma <- 0.572612

# Step 1.3: Create the lognormal kernel matrix
kernel_matrix_LH <- dlnorm(distance_matrix_LH, meanlog = mu, sdlog = sigma)

# Step 1.4: Check the dimensions and a small portion of the matrix
dim(kernel_matrix_LH)
print("Top-left corner of the matrix:")
print(kernel_matrix_LH[1:10, 1:10])  # Displaying the top-left 10x10 cells

#Step 2: create the population matrix
#Step 2.1: create an identifcal raster and print basic information about the raster
raster_1km_human_population_threshold_3 <- raster_1km_human_population_threshold
print(raster_1km_human_population_threshold_3)
# Summary of the raster to check the range of values
summary(raster_1km_human_population_threshold_3)
#Step 2.2: Check the number of non-zero and NA values
num_zeros <- sum(values(raster_1km_human_population_threshold_3) == 0, na.rm = TRUE)
num_nas <- sum(is.na(values(raster_1km_human_population_threshold_3)))
num_non_zeros <- sum(values(raster_1km_human_population_threshold_3) > 0, na.rm = TRUE)
cat("Number of zero values:", num_zeros, "\n")
cat("Number of NA values:", num_nas, "\n")
cat("Number of non-zero values:", num_non_zeros, "\n")
#Step 2.3: Replace NA values with zeros (if appropriate for your analysis)
raster_1km_human_population_threshold_3[is.na(raster_1km_human_population_threshold_3)] <- 0
plot(raster_1km_human_population_threshold_3)
# Check if NA values are replaced
na_count_after <- sum(is.na(values(raster_1km_human_population_threshold_3)))
print(paste("Number of NA values after replacement:", na_count_after))
#Step 2.4: Get the number of rows and columns from the original raster
nrows <- nrow(raster_1km_human_population_threshold_3)
ncols <- ncol(raster_1km_human_population_threshold_3)
#Step 2.5: Convert the raster to a matrix and reshape it
population_matrix_LH <- matrix(values(raster_1km_human_population_threshold_3), 
                            nrow = nrows, ncol = ncols, byrow = TRUE)
# Check the dimensions and a small portion of the matrix
dim(population_matrix_LH)
print("Top-left corner of the matrix:")
print(population_matrix_LH[1:10, 1:10])  # Displaying the top-left 10x10 cells

# Step 3: Use the 'Fast-Fourier' transform to compute the convolution
convolved_matrix_LH <- convolve_image(population_matrix_LH, kernel_matrix_LH)

# Step 4: convert the matrix to raster
# Step 4.1: Convert the convolved matrix back to raster format using terra's rast function
convolved_raster_LH <- rast(nrows = nrow(raster_1km_human_population_threshold_3),
                                  ncols = ncol(raster_1km_human_population_threshold_3),
                                  xmin = xmin(raster_1km_human_population_threshold_3),
                                  xmax = xmax(raster_1km_human_population_threshold_3),
                                  ymin = ymin(raster_1km_human_population_threshold_3),
                                  ymax = ymax(raster_1km_human_population_threshold_3))
# Step 4.2: Assign the values of the convolved matrix back to the raster
values(convolved_raster_LH) <- convolved_matrix_LH
# Step 4.3: fix crs, extent and resolution
res(raster_1km_human_population_threshold_3)
res(convolved_raster_LH)
crs(raster_1km_human_population_threshold_3)
crs(convolved_raster_LH)
ext(raster_1km_human_population_threshold_3)
ext(convolved_raster_LH)
# Step 4.3: fix crs
crs(convolved_raster_LH) <- crs(raster_1km_human_population_threshold_3)

# Step 5: Plot results
plot(raster_1km_human_population_threshold_3, main = "Original Population Raster")
plot(kernel_matrix_LH, main = "Kernel Raster")
plot(convolved_raster_LH, main = "Convolved Raster with Identity Kernel")

# Step 6: mask the ocean
# Step 6.1: import the inland shapefile
inland_shapefile <- st_read("S:/Magali_Weissgerber/Europe frontiers_scratch/NUT3_from_rewilding_folder/Europe_limit_from_nut3_dissolve.shp")
# Step 6.2: mask the output raster
convolved_raster_masked_LH <- mask(convolved_raster_LH, inland_raster)
plot(convolved_raster_masked_LH)

# Compute quantile breaks (exclude NA values)
q_breaks <- quantile(values(convolved_raster_masked_LH), 
                     probs = seq(0, 1, length.out = 11), 
                     na.rm = TRUE)

# Plot with 10 quantile levels
plot(convolved_raster_masked_LH, 
     breaks = q_breaks, 
     col = hcl.colors(10, "YlOrRd", rev = TRUE), 
     main = "Convolved Raster (Quantile Breaks) large herbivores")

#Step 7: export
#output_file_WTT_masked_LH <- ("S:/Magali_Weissgerber/Species raster for social benefit of rewilding/WTT_LH_raster_masked_241212.tif")
#writeRaster(convolved_raster_masked_LH, output_file_WTT_masked_LH, overwrite=TRUE)

```

## for each cell multiply the wildlife status (current presence where cell value = 1 and comeback probability where cell value is between 0 and 1) with the sum of people travelling

```{r}

#Get the occupancy raster (with 1, 0 and NA)
raster_1km_wildlife_occupancy_stack_path <- "S:/Magali_Weissgerber/Species raster for social benefit of rewilding/SP_occupancy_250106.tif" # Read the 1km resolution raster
raster_1km_wildlife_occupancy_stack <- rast(raster_1km_wildlife_occupancy_stack_path)
plot(raster_1km_wildlife_occupancy_stack)

#Get the probability based on the distance to 1 (cells with 1 have a 0 probability)
raster_1km_wildlife_probability_stack_path <- "S:/Magali_Weissgerber/Species raster for social benefit of rewilding/SP_expansion_probability_250106.tif" # Read the 1km resolution raster
raster_1km_wildlife_probability_stack <- rast(raster_1km_wildlife_probability_stack_path)
plot(raster_1km_wildlife_probability_stack)

#WTT changed with the review, hence the files WTT_LC_raster_masked_240925.tif and WTT_LH_raster_masked_241212.tif are not valid anymore but convolved_raster_masked_LC and convolved_raster_masked_LH are produced above the right new way
#Get WTT
#Carnivores
#convolved_raster_masked_LC <- "S:/Magali_Weissgerber/Species raster for social benefit of rewilding/WTT_LC_raster_masked_240925.tif" # Read raster
#convolved_raster_masked_LC <- rast(convolved_raster_masked_LC)
plot(convolved_raster_masked_LC)
#Herbivores
#convolved_raster_masked_LH <- "S:/Magali_Weissgerber/Species raster for social benefit of rewilding/WTT_LH_raster_masked_241212.tif" # Read raster
#convolved_raster_masked_LH <- rast(convolved_raster_masked_LH)
plot(convolved_raster_masked_LH)

#Multiply occupancy x WTT and multiply probability x WTT

# Define species for carnivores and herbivores (without prefixes)
carnivores <- c("CANLUP", "LYNLYN", "LYNPAR", "URSARC")  # Example carnivores
herbivores <- c("ALCALC", "BISBON", "CAPIBE", "CAPPYR")  # Example herbivores

# Construct species names for occupancy (M_ prefix)
carnivores_occupancy <- paste0("M_", carnivores)
herbivores_occupancy <- paste0("M_", herbivores)
# Construct species names for probability (P_ prefix)
carnivores_probability <- paste0("P_", carnivores)
herbivores_probability <- paste0("P_", herbivores)

# Identify indices in the stacks
# Occupancy stack
carnivore_indices_occupancy <- which(names(raster_1km_wildlife_occupancy_stack) %in% carnivores_occupancy)
herbivore_indices_occupancy <- which(names(raster_1km_wildlife_occupancy_stack) %in% herbivores_occupancy)
# Probability stack
carnivore_indices_probability <- which(names(raster_1km_wildlife_probability_stack) %in% carnivores_probability)
herbivore_indices_probability <- which(names(raster_1km_wildlife_probability_stack) %in% herbivores_probability)

# Split the stacks by species type
# For occupancy
occupancy_carnivores <- raster_1km_wildlife_occupancy_stack[[carnivore_indices_occupancy]]
occupancy_herbivores <- raster_1km_wildlife_occupancy_stack[[herbivore_indices_occupancy]]
# For probability
probability_carnivores <- raster_1km_wildlife_probability_stack[[carnivore_indices_probability]]
probability_herbivores <- raster_1km_wildlife_probability_stack[[herbivore_indices_probability]]

# Multiply rasters
SB_occupancy_carnivores <- occupancy_carnivores * convolved_raster_masked_LC
SB_occupancy_herbivores <- occupancy_herbivores * convolved_raster_masked_LH
SB_probability_carnivores <- probability_carnivores * convolved_raster_masked_LC
SB_probability_herbivores <- probability_herbivores * convolved_raster_masked_LH

# Combine the results back into stacks
SB_occupancy <- c(SB_occupancy_carnivores, SB_occupancy_herbivores)
SB_probability <- c(SB_probability_carnivores, SB_probability_herbivores)

# Rename layers to indicate SB_occupancy and SB_probability
names(SB_occupancy) <- gsub("^M_", "SB_occ_", names(SB_occupancy))
names(SB_probability) <- gsub("^P_", "SB_proba_", names(SB_probability))

# Calculate the sum of all layers in each stack
SB_occ_sum <- sum(SB_occupancy)  # Sum of all SB_occ rasters
SB_proba_sum <- sum(SB_probability)  # Sum of all SB_proba rasters

# Rename the sum rasters for clarity
names(SB_occ_sum) <- "SB_occ_sum"
names(SB_proba_sum) <- "SB_proba_sum"

# Add the sum raster to each stack
SB_occupancy <- c(SB_occupancy, SB_occ_sum)
SB_probability <- c(SB_probability, SB_proba_sum)

# Plot
plot(SB_occupancy)
plot(SB_probability)

# Verify the new order
print(names(SB_occupancy))  # Should match desired_order_occupancy
# Re-order SB_occupancy and SB_probability
desired_order_occupancy <- c(
  "SB_occ_CANLUP", "SB_occ_ALCALC", "SB_occ_BISBON", "SB_occ_LYNLYN", 
  "SB_occ_URSARC", "SB_occ_LYNPAR", "SB_occ_CAPIBE", "SB_occ_CAPPYR", 
  "SB_occ_sum"
)

desired_order_probability <- c(
  "SB_proba_CANLUP", "SB_proba_ALCALC", "SB_proba_BISBON", "SB_proba_LYNLYN", 
  "SB_proba_URSARC", "SB_proba_LYNPAR", "SB_proba_CAPIBE", "SB_proba_CAPPYR", 
  "SB_proba_sum"
)

# Reorder SB_occupancy
SB_occupancy <- SB_occupancy[[match(desired_order_occupancy, names(SB_occupancy))]]

# Reorder SB_probability
print(names(SB_occupancy))  # Should match desired_order_occupancy
print(names(SB_probability))  # Should match desired_order_probability

#Stack WTT, occupancy, proba and the two SB
WTT_occ_proba_SB_stack <- c(convolved_raster_masked_LC, convolved_raster_masked_LH, raster_1km_wildlife_occupancy_stack, raster_1km_wildlife_probability_stack, SB_occupancy, SB_probability)
names(WTT_occ_proba_SB_stack)[1:2] <- c("WTT_LC", "WTT_LH")

print(names(WTT_occ_proba_SB_stack)) 

#Export
#output_WTT_occ_proba_SB_stack <- ("S:/Magali_Weissgerber/Species raster for social benefit of rewilding/WTT_occ_proba_SB_stack_250106.tif")
#writeRaster(WTT_occ_proba_SB_stack, output_WTT_occ_proba_SB_stack, overwrite=TRUE)

WTT_occ_proba_SB_stack <- c(convolved_raster_masked_LC, convolved_raster_masked_LH, raster_1km_wildlife_occupancy_stack, raster_1km_wildlife_probability_stack)
names(WTT_occ_proba_SB_stack)[1:2] <- c("WTT_LC", "WTT_LH")
plot(WTT_occ_proba_SB_stack)

#output_WTT_occ_proba_SB_stack <- ("S:/Magali_Weissgerber/Species raster for social benefit of rewilding/WTT_occ_proba_SB_stack_250912.tif")
#writeRaster(WTT_occ_proba_SB_stack, output_WTT_occ_proba_SB_stack, overwrite=TRUE)

```

## Step 3 - Determining social conflict

1) Import livestock distribution data

```{r}




```

## Step n

```{r}




```

## Step n

```{r}




```

## Step n

```{r}




```

## Step n

```{r}




```

## Step n

```{r}




```

## Step n

```{r}




```

## Step n

```{r}




```
